# Introduce a trader budget and calculating the quantity

## Objectives

- fetch step_size
- append budget and step_size to the `Trader`'s state compiled by the `Leader`
- append budget and step_size to the `Trader`'s state
- calculate quantity

## Fetch `step_size`

In the 2nd chapter we hardcoded `quantity` to 100, it's time to refactor that. We will need `step_size` information from the Binance which we are
already retrieving together with `tick_size` in the `exchangeInfo` call(but not getting it out from the response). So we will rename the `fetch_tick_size/1` function to `fetch_symbol_filters/1` which will allow us to return multiple filters(`tick_size` and `step_size`) from that function.

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/leader.ex
  ...
  defp fetch_symbol_settings(symbol) do
    symbol_filters = fetch_symbol_filters(symbol) # <= updated fetch_tick_size
    
    Map.merge(
      %{
        symbol: symbol,
        chunks: 1,
        budget: 20,
        # -0.01% for quick testing
        buy_down_interval: "0.0001",
        # -0.12% for quick testing
        profit_interval: "-0.0012"
      },
      symbol_filters
    )
  end

  defp fetch_symbol_filters(symbol) do  # <= updated fetch_tick_size
    symbol_filters =
      @binance_client.get_exchange_info()
      |> elem(1)
      |> Map.get(:symbols)
      |> Enum.find(&(&1["symbol"] == symbol))
      |> Map.get("filters")

    tick_size =
      symbol_filters
      |> Enum.find(&(&1["filterType"] == "PRICE_FILTER"))
      |> Map.get("tickSize")

    step_size =
      symbol_filters
      |> Enum.find(&(&1["filterType"] == "LOT_SIZE"))
      |> Map.get("stepSize")

    %{
      tick_size: tick_size,
      step_size: step_size
    }
  end
```

Instead of reassigning the filters one by one into the settings, we will merge them together(#1). Additionally, we will introduce a `budget`(#2) which will be shared across all traders of the symbol. Also, we don't need to assign `tick_size` here as it's part of the settings that are merged.

## Append `budget` and `step_size` to the `Trader`'s state inside the `Leader` 

The `budget` needs to be added to the `%State{}`(`step_size` will be automatically passed on by `struct/2`) of the trader inside `fresh_trader_state/1`(where we initialize the state of traders). Before we will assign it we need to divide it by the number of chunks as each trader gets only a chunk of the budget:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/leader.ex
  defp fresh_trader_state(settings) do
    %{
      struct(Trader.State, settings) |
      budget: D.div(settings.budget, settings.chunks)
    }
  end
```

In the code above we are using the `Decimal` module(aliased as `D`) to calculate the budget - we need to alias it at the top of `Naive.Leader`'s file: 

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/leader.ex
defmodule Naive.Leader do
  use GenServer

  alias Decimal, as: D # <= add this line
  alias Naive.Trader
  ...
```

## Append `budget` and `step_size` to the `Trader`'s state

We need to add both `budget` and `step_size` to the `Naive.Trader`'s state struct:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/trader.ex
  ...
  defmodule State do
    @enforce_keys [
      :symbol,
      :budget, # <= add this line
      :buy_down_interval,
      :profit_interval,
      :tick_size,
      :step_size # <= add this line and comma above
    ]
    defstruct [
      :symbol,
      :budget, # <= add this line
      :buy_order,
      :sell_order,
      :buy_down_interval,
      :profit_interval,
      :tick_size,
      :step_size # <= add this line and comma above
    ]
  end
  ...
```

## Calculate quantity

Jumping back to the `handle_info/2` where the `Naive.Trader` places a buy order, we need to pattern match on the `step_size` and `budget` then we will be able to swap hardcoded quantity with the result of calling the `calculate_quantity/3` function:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/trader.ex
  ...
  def handle_info(
        %TradeEvent{price: price},
        %State{
          symbol: symbol,
          budget: budget, # <= add this line
          buy_order: nil,
          buy_down_interval: buy_down_interval,
          tick_size: tick_size,
          step_size: step_size # <= add this line
        } = state
      ) do
    ...
    quantity = calculate_quantity(budget, price, step_size)
    ...
```

To calculate quantity we will just divide the `budget` by the `price` with a caveat that it's possible (as with calculating the price) that it's not a legal quantity value as it needs to be divisible by `step_size`:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/trader.ex
  # add below at the bottom of the file
  ...
  defp calculate_quantity(budget, price, step_size) do
    # not necessarily legal quantity
    exact_target_quantity = D.div(budget, price)

    D.to_string(
      D.mult(
        D.div_int(exact_target_quantity, step_size),
        step_size
      ),
      :normal
    )
  end
```

### IEx testing

That finishes the `quantity`(and `budget`) implementation, we will jump into the IEx session to see how it works.

First, start the streaming and trading on the same symbol and a moment later you should see a variable amount of quantity that more or less uses the full allowed budget:

```{r, engine = 'bash', eval = FALSE}
$ iex -S mix
...
iex(1)> Streamer.start_streaming("XRPUSDT")
{:ok, #PID<0.313.0>}
iex(2)> Naive.start_trading("XRPUSDT")
21:16:14.829 [info]  Starting new supervision tree to trade on XRPUSDT
21:16:16.755 [info]  Initializing new trader for XRPUSDT
21:16:20.009 [info]  Placing BUY order for XRPUSDT @ 0.29506, quantity: 67.7
21:16:23.456 [info]  Buy order filled, placing SELL order for XRPUSDT @ 0.29529,
quantity: 67.7
```

As we can see our `Naive.Trader` process is now buying and selling based on passed budget.

[Note] Please remember to run the `mix format` to keep things nice and tidy.

Source code for this chapter can be found at [Github](https://github.com/frathon/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_07)
