# Introduce a trader budget and calculating the quantity

In the previous chapter, we made our traders smarter about *when* to buy by introducing the `buy_down_interval`.
But there's still a glaring problem: we've been hardcoding the quantity to 100 units every single trade.

That's problematic for two reasons. First, 100 units of XRPUSDT (around $160) is very different from 100 units of BTCUSDT (around $9 million!).
Second, when we start running multiple traders in parallel, we'll need to divide our budget between them -
and we can't do that if the quantity is hardcoded.

In this chapter, we'll introduce a proper `budget` for each symbol and calculate the `quantity` dynamically based on that budget and the current price.
We'll also fetch the `step_size` from Binance, which tells us the minimum quantity increment for each symbol (just like `tick_size` does for prices).

Let's make our traders budget-aware!

## Objectives

- fetch `step_size`
- append `budget` and `step_size` to the `Naive.Trader`'s state inside the `Naive.Leader`
- add `budget` and `step_size` to the `Naive.Trader`'s `State` struct
- calculate `quantity`

## Fetch `step_size`

In the second chapter we hardcoded `quantity` to 100 - it's time to refactor that. Just like prices must be divisible by `tick_size`,
quantities must be divisible by `step_size` - Binance will reject orders with invalid quantities.

We're already retrieving this information in the `exchangeInfo` call alongside `tick_size`, but we're not extracting it from the response yet.

So we will rename the `fetch_tick_size/1` function to `fetch_symbol_filters/1`
which will allow us to return multiple filters(`tick_size` and `step_size`) from that function.

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/leader.ex
  ...
  defp fetch_symbol_settings(symbol) do
    symbol_filters = fetch_symbol_filters(symbol) # <= updated fetch_tick_size
    
    Map.merge(
      %{
        symbol: symbol,
        chunks: 1,
        budget: 20,
        # -0.01% for quick testing
        buy_down_interval: "0.0001",
        # -0.12% for quick testing
        profit_target: "-0.0012"
      },
      symbol_filters
    )
  end

  defp fetch_symbol_filters(symbol) do  # <= updated fetch_tick_size
    {:ok, exchange_info} = @binance_client.get_exchange_info()

    symbol_filters =
      exchange_info
      |> Map.get(:symbols)
      |> Enum.find(&(&1["symbol"] == symbol))
      |> Map.get("filters")

    tick_size =
      symbol_filters
      |> Enum.find(&(&1["filterType"] == "PRICE_FILTER"))
      |> Map.get("tickSize")

    step_size =
      symbol_filters
      |> Enum.find(&(&1["filterType"] == "LOT_SIZE"))
      |> Map.get("stepSize")

    %{
      tick_size: tick_size,
      step_size: step_size
    }
  end
```

Instead of reassigning the filters one by one into the settings, we merge them together using `Map.merge/2`.
We also introduce a `budget` which will be shared across all traders of the symbol.
Note that we don't need to explicitly assign `tick_size` anymore - it's included in `symbol_filters` and gets merged automatically.

## Append `budget` and `step_size` to the `Naive.Trader`'s state inside the `Naive.Leader`

The `budget` needs to be calculated and added to the trader's `%State{}` inside `fresh_trader_state/1`.
Since each trader gets only a chunk of the total budget, we divide it by the number of chunks before assigning.
The `step_size`, on the other hand, passes through automatically - `struct/2` copies any matching keys from the settings map into the struct:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/leader.ex
  defp fresh_trader_state(settings) do
    %{
      struct(Trader.State, settings)
      | budget: D.div(settings.budget, settings.chunks)
    }
  end
```

In the code above we are using the `Decimal` module(aliased as `D`) to calculate the budget - we need to alias it at the top of `Naive.Leader`'s file:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/leader.ex
defmodule Naive.Leader do
  use GenServer

  alias Decimal, as: D # <= add this line
  alias Naive.Trader
  ...
```

## Add `budget` and `step_size` to the `Naive.Trader`'s `State` struct

We need to add both `budget` and `step_size` to the `Naive.Trader`'s state struct:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/trader.ex
  ...
  defmodule State do
    @enforce_keys [
      :symbol,
      :budget, # <= add this line
      :buy_down_interval,
      :profit_target,
      :tick_size,
      :step_size # <= add this line and comma above
    ]
    defstruct [
      :symbol,
      :budget, # <= add this line
      :buy_order,
      :sell_order,
      :buy_down_interval,
      :profit_target,
      :tick_size,
      :step_size # <= add this line and comma above
    ]
  end
  ...
```

## Calculate quantity

Jumping back to the `handle_info/2` where the `Naive.Trader` places a buy order,
we need to pattern match on the `step_size` and `budget`,
then we will be able to swap the hardcoded quantity with the result of calling the `calculate_quantity/3` function:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/trader.ex
  ...
  def handle_info(
        %TradeEvent{price: price},
        %State{
          symbol: symbol,
          budget: budget, # <= add this line
          buy_order: nil,
          buy_down_interval: buy_down_interval,
          tick_size: tick_size,
          step_size: step_size # <= add this line
        } = state
      ) do
    ...
    quantity = calculate_quantity(budget, price, step_size)
    ...
```

To calculate quantity, the formula is simple: `quantity = budget / price`. For example, with a $20 budget and a price of $0.30, we'd buy about 66.67 units.
But there's a caveat: just like prices must align to `tick_size`, quantities must align to `step_size`.
So we'll round down to the nearest valid quantity:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/lib/naive/trader.ex
  # add below at the bottom of the file
  ...
  defp calculate_quantity(budget, price, step_size) do
    # Round down to nearest valid step_size
    # Example:
    #   target quantity = 67.86
    #   step_size = 0.1
    #   67.86 / 0.1 = 678.6 → truncate to 678 → 678 * 0.1 = 67.8
    budget
    |> D.div(D.from_float(price))
    |> D.div_int(step_size)
    |> D.mult(step_size)
    |> D.to_string(:normal)
  end
```

### Manual testing in IEx

That finishes the `quantity`(and `budget`) implementation, we will jump into the IEx session to see how it works.

First, start the streaming and trading on the same symbol.
A moment later you should see the quantity being calculated from the budget rather than hardcoded:

```{r, engine = 'bash', eval = FALSE}
$ iex -S mix
...
iex(1)> Streamer.start_streaming("XRPUSDT")
{:ok, #PID<0.313.0>}
iex(2)> Naive.start_trading("XRPUSDT")
21:16:14.829 [info]  Starting new supervision tree to trade on XRPUSDT
21:16:16.755 [info]  Initializing new trader for XRPUSDT
21:16:20.009 [info]  Placing BUY order for XRPUSDT @ 0.29506, quantity: 67.7
21:16:23.456 [info]  Buy order filled, placing SELL order for XRPUSDT @ 0.29529,
quantity: 67.7
```

As we can see our `Naive.Trader` process is now buying and selling based on the passed budget.
The quantity of 67.7 comes from dividing our $20 budget by the price of ~$0.295.

**Great progress!** Our traders now:

- Calculate quantity dynamically based on an assigned budget
- Respect Binance's `step_size` requirement for valid quantities
- Are ready to share a total budget across multiple parallel traders (via the `chunks` setting)

**So, What's Next?**

We have all the pieces in place for something more ambitious. Right now, we start exactly one trader per symbol,
and when it finishes a trade cycle, a new one starts. But what if the price keeps dropping after we buy?
We're sitting there holding our position, watching opportunities pass by.

In the next chapter, we'll implement a "rebuy" mechanism. When the price drops below our buy price by a certain interval,
we'll spin up another trader to buy more at the lower price. This lets us scale into positions as the price falls
and potentially profit more when it recovers. We'll finally put that `chunks` setting to work!

[Note] Please remember to run `mix format` to keep things nice and tidy.

The source code for this chapter can be found in the book's source code repository
(branch:
[chapter_07](https://github.com/Cinderella-Man/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_07)).
