# Mox rocks

In the last chapter, we built an end-to-end test that broadcasts trade events
and verifies orders land in the database.
It works - but it's slow, brittle, and when it fails, we're left guessing
where the problem is.

Here's the thing about end-to-end tests: they're great for proving the system
works as a whole, but terrible for developer productivity. Every test run
requires database setup, process orchestration, and hardcoded sleeps.
And when something breaks, good luck figuring out which module caused it.

What we need is fast, focused unit tests. We'll use the `mox` library to mock
the dependencies of our `Naive.Trader` - PubSub, Binance, the Leader, even Logger.
Each mock will pattern match on expected values, giving us assertions for free.
The result? Tests that run in milliseconds, pinpoint failures exactly, and
don't need a database in sight.

Along the way, we'll define behaviours for modules that don't have them,
wire up environment-based configuration, and see why `mox`'s approach to
mocking is fundamentally better than hand-rolled stubs.

## Objectives
- introduction to mock-based tests
- add the Mox package
- identify the `Naive.Trader` dependencies
  * mock the `Binance` module
  * mock the `NaiveLeader` module
  * mock the `Phoenix.PubSub` module
  * mock the `Logger` module
- implement a test of the `Naive.Trader` module
- define an alias to run unit tests

## Introduction to mock-based tests

Let's revisit the specific downsides of our end-to-end test from the last chapter:

- we will be unable to run more than one end-to-end test in parallel as they rely on the database's state
- we need to set up the database before every test run
- we need to start processes in the correct order with the suitable parameters
- we need to wait a (guessed) hardcoded amount of time that it will take to finish the trading(this is extremely bad as it will cause randomly failing tests as people will make the time shorter to speed up tests)
- we wouldn't be able to quickly pinpoint which part the error originated from as the test spans over a vast amount of the system
- logging was polluting our test output

How could we fix the above issues?

The most common way is to limit the scope of the test.
Instead of testing the whole trading flow, we could focus on testing a single `Naive.Trader` process.

Focusing on a single trader process would remove the requirement for starting multiple processes before testing,
but it would also bring its own challenges.

Let's look at a concrete example:

When the `Naive.Trader` process starts, it subscribes to the `TRADE_EVENTS:#{symbol}` PubSub topic.
It also broadcasts updates of the orders it placed to the `ORDERS:#{symbol}` PubSub topic.

How could we break the link between the `Naive.Trader` and the PubSub(or any other module it depends on)?

We could utilize the trick that we used for the `Binance` module. We could create a module that provides the same functions as the `PubSub` module.

We know that the trader process calls `Phoenix.PubSub.subscribe/2` and `Phoenix.PubSub.broadcast/3` functions.
We could implement a module that contains the same functions:

```{r, engine = 'elixir', eval = FALSE}
defmodule Test.PubSub do
    def subscribe(_, _), do: :ok
    def broadcast(_, _, _), do: :ok
end
```

The above module would satisfy the PubSub's functionality required by the `Naive.Trader` module, but this solution comes with a couple of drawbacks:

- it doesn't test the passed values. It just ignores them, which is a missed opportunity to confirm that the PubSub module was called with the expected values
- we can't define a custom implementation specific to a test. This is possible by extending the implementation with test-related returns(hack!)

Using the `mox` module would fix both of the problems mentioned above. With the `mox` module we can define ad-hoc function implementation per test:

```{r, engine = 'elixir', eval = FALSE}
    # inside test file
    test ...
        Test.PubSubMock
        |> expect(:subscribe, fn (_module, "TRADE_EVENTS:XRPUSDT") -> :ok end)
        |> expect(:broadcast, fn (_module, "ORDERS:XRPUSDT", _order) -> :ok end)
```

There are multiple benefits to using the `mox` module instead of handcrafting the implementation:

- it allows defining functions that will pattern match values specific to each test(as in the case of the "usual" pattern matching, they will break when called with unexpected values)
- it allows defining implementations of the mocked functions based on incoming(test-specific) values
- it can validate that all defined mocked functions have been called by the test
- it comes with its own tests, so we don't need to test the mocking framework itself - unlike a custom handcrafted implementation that we'd have to maintain and verify

But there's a catch ;)

For the `mox` to know what sort of functions the module provides, it needs to know its `behaviour`.

In Elixir, to define a behaviour of the module, we need to add the `@callback` attributes to it:

```{r, engine = 'elixir', eval = FALSE}
  defmodule Core.Test.PubSub do
    @type t :: atom
    @type topic :: binary
    @type message :: term

    @callback subscribe(t, topic) :: :ok | {:error, term}
    @callback broadcast(t, topic, message) :: :ok | {:error, term}
  end
```

A `behaviour` can be defined in a separate module if we are working with a 3rd party module
that doesn't provide it(like in the case of the `Phoenix.PubSub` module).

Note: The additional benefit of using the `behaviours` is that we could tell Elixir that our module *implements* the behaviour
by adding the `@behaviour` attribute:

```{r, engine = 'elixir', eval = FALSE}
defmodule MyPubSub do
    @behaviour Core.Test.PubSub
    ...
```

Using the above will cause Elixir to validate at compile time that the `MyPubSub` module
implements all functions defined inside the `Core.Test.PubSub` module(otherwise it will raise a compilation error).

Let's get back to the main topic. We figured out that we could mock all of the modules that the `Naive.Trader` depends on using the `mox` module. 

But, how would we tell the `Naive.Trader` to use the mocked modules instead of the "real" ones when we run tests?

We could make all modules that the `Naive.Trader` depends on be dynamically injected from the configuration(based on the environment).

Another requirement to make `mox` work is to define the mocks upfront using the `Mox.defmock/2` function
(called in `test_helper.exs`, which runs before tests).
It will dynamically define a new module that will be limited by the passed behaviour -
we will only be able to mock functions defined as part of that behaviour inside our tests.

To sum up, there are a few steps to get the `mox` running:

- implement behaviours that we would like to mock(as most packages [like `Phoenix.PubSub`] don't include them)
- define mock modules using the `Mox.defmock` function
- modify the application's configuration to use the mocked module(s)
- specify mocked module's expectation inside the test

That's the full mental model. Now let's see how each step plays out in practice.

## Add the `mox` package

First let's add the `mox` package to the `naive` application's dependencies:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/mix.exs
  ...
  defp deps do
    [
      ...
      {:mox, "~> 1.0", only: [:test, :integration]},
      ...
```

We can now run `mix deps.get` to fetch the `mox` package.

[Note] As we will add the `mox`'s mocking code into the `test_helper.exs` file, we need to
make `mox` available in all test environments(both `test` and `integration`).

## Identify the `Naive.Trader` dependencies

Let's investigate the `Naive.Trader` module(`/apps/naive/lib/naive/trader.ex`). We are looking for all calls to other modules - we can see:

- `Logger.info/1`
- `Phoenix.PubSub.subscribe/2`
- `@binance_client.order_limit_buy/4`
- `Naive.Leader.notify/2`
- `@binance_client.get_order/3`
- `@binance_client.order_limit_sell/4`
- `Phoenix.PubSub.broadcast/3`

So the `Naive.Trader` relies on four modules:

- `Logger`
- `Phoenix.PubSub`
- `Naive.Leader`
- `@binance_client`(either `Binance` or `BinanceMock`)

We will need to work through them one by one.

### Mock the `Binance` module

Let's start with the binance client, as it's already a dynamic value based on the configuration.

Neither the `Binance` nor the `BinanceMock`(our dummy implementation) module provides a behaviour -
let's fix that by defining the `@callback` attributes at the top of the `BinanceMock` module before the structs:

```{r, engine = 'elixir', eval = FALSE}
# /apps/binance_mock/lib/binance_mock.ex
  ...
  alias Binance.Order
  alias Binance.OrderResponse
  alias Core.Struct.TradeEvent

  @type symbol :: binary
  @type quantity :: binary
  @type price :: binary
  @type time_in_force :: binary
  @type timestamp :: non_neg_integer
  @type order_id :: non_neg_integer
  @type orig_client_order_id :: binary
  @type recv_window :: binary

  @callback order_limit_buy(
              symbol,
              quantity,
              price,
              time_in_force
            ) :: {:ok, %OrderResponse{}} | {:error, term}

  @callback order_limit_sell(
              symbol,
              quantity,
              price,
              time_in_force
            ) :: {:ok, %OrderResponse{}} | {:error, term}

  @callback get_order(
              symbol,
              timestamp,
              order_id,
              orig_client_order_id | nil,
              recv_window | nil
            ) :: {:ok, %Order{}} | {:error, term}
```

In the above code, we added three `@callback` attributes that define the binance client behaviour.
For clarity, we defined a distinct type for each of the arguments.

As we now have a binance client behaviour defined,
we can use it to define a mock using the `Mox.defmock/2` function inside the `test_helper.exs` file of the `naive` application:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/test_helper.exs
ExUnit.start()

Application.ensure_all_started(:mox) #1

Mox.defmock(Test.BinanceMock, for: BinanceMock) #2
```

First(#1), we need to ensure that the `mox` application has been started. Then(#2),
we can tell the `mox` package to define the `Test.BinanceMock` module based on the `BinanceMock` behaviour.

As we defined the binance client behaviour and mock, we can update our configuration to use them.
We want to keep using the `BinanceMock` module in the development environment, but for the `test` environment,
we would like to set the mocked module generated by the `mox` package:

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
config :naive,
  binance_client: Test.BinanceMock
``` 

### Mock the `NaiveLeader` module

We can now move back to the `Naive.Trader` module to update all the hardcoded references to the `Naive.Leader` module
with a dynamic attribute called `@leader` and add this attribute at the top of the module:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/trader.ex
  ...
  @leader Application.compile_env(:naive, :leader)
  ...
  @leader.notify(:trader_state_updated, new_state)
  ...
  @leader.notify(:trader_state_updated, new_state)
  ...
  @leader.notify(:rebuy_triggered, new_state)
  ...
```

As with the `BinanceMock`(our dummy implementation) module, the `Naive.Leader` module doesn't provide a behaviour -
let's fix that by defining the `@callback` attributes at the top of the module:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/leader.ex
  ...
  @type event_type :: atom
  @callback notify(event_type, %Trader.State{}) :: :ok
```

In the above code, we added a single `@callback` attribute that defines the naive leader behaviour.
For clarity, we defined a distinct type for the `event_type` argument.

As we now have a naive leader behaviour defined, we can use it to define a mock using the `Mox.defmock/2` function
inside the `test_helper.exs` file of the `naive` application:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/test_helper.exs
Mox.defmock(Test.Naive.LeaderMock, for: Naive.Leader)
``` 

In the above code, we've told the `mox` package to define the `Test.Naive.LeaderMock` module based on the `Naive.Leader` behaviour.

We are moving on to the configuration.
As the `Naive.Leader` wasn't part of the configuration, we need to add it to the default config and test config files.

First, let's add the `:leader` key inside the `config :naive` in the default `/config/config.exs` configuration file:

```{r, engine = 'elixir', eval = FALSE}
# /config/config.exs
...
config :naive,
  binance_client: BinanceMock,
  leader: Naive.Leader,        # <= added
  ...
``` 

and then we need to apply the same update to the `/config/test.exs` configuration file
(it will point to the module generated by the `mox` package - `Test.Naive.LeaderMock`):

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
...
config :naive,
  binance_client: Test.BinanceMock,
  leader: Test.Naive.LeaderMock     # <= added
  ...
``` 



### Mock the `Phoenix.PubSub` module

Mocking the `Phoenix.PubSub` dependency inside the `Naive.Trader` module will look very similar to the last two mocked modules.



Inside the `Naive.Trader` module we need to update all the references to the `Phoenix.PubSub` to an `@pubsub_client` attribute
with value dependent on the configuration:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/trader.ex
  ...
  @pubsub_client Application.compile_env(:core, :pubsub_client)
  ...
    @pubsub_client.subscribe(
      Core.PubSub,
      "TRADE_EVENTS:#{symbol}"
    )
  ...
    @pubsub_client.broadcast(
      Core.PubSub,
      "ORDERS:#{order.symbol}",
      order
    )
  ...
```

The `Phoenix.PubSub` module doesn't provide a behaviour. As we can't modify its source,
we need to create a new module to define the PubSub behaviour.
This is the same `Core.Test.PubSub` module we used as an example earlier - now we'll actually create it.
Let's add a new file called `test.ex` inside the `/apps/core/lib/core` directory with the following behaviour definition:

```{r, engine = 'elixir', eval = FALSE}
# /apps/core/lib/core/test.ex
defmodule Core.Test do
  defmodule PubSub do
    @type t :: atom
    @type topic :: binary
    @type message :: term

    @callback subscribe(t, topic) :: :ok | {:error, term}
    @callback broadcast(t, topic, message) :: :ok | {:error, term}
  end
end
```

As previously, we defined a couple of callbacks and additional types for each of their arguments.



Next, we will use the above behaviour to define a mock using the `Mox.defmock/2` function inside the `test_helper.exs` file of the `naive` application:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/test_helper.exs
Mox.defmock(Test.PubSubMock, for: Core.Test.PubSub)
``` 

In the above code, we've told the `mox` package to define the `Test.PubSubMock` module based on the `Core.Test.PubSub` behaviour.

The final step will be to append the `:core, :pubsub_client` configuration to the `/config/config.exs` file:

```{r, engine = 'elixir', eval = FALSE}
# /config/config.exs
config :core,                  # <= added
  pubsub_client: Phoenix.PubSub # <= added
``` 

and the test `/config/test.exs` configuration file:

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
config :core,                  # <= added
  pubsub_client: Test.PubSubMock # <= added
```

### Mock the `Logger` module

Before we dive in, we should ask ourselves: why mock the `Logger` module?

We could raise the logging level to `error` and be done with it.
Yes, that would fix all debug/info/warning logs,
but we would also miss an opportunity to confirm a few details (depends on what's necessary for our use case):

- you can ensure that the log was called when the tested function was run
- you can pattern match the logging level
- you can check the message. This could be useful if you don't want to put sensitive information like banking details etc. inside log messages

Mocking the `Logger` dependency inside the `Naive.Trader` module will follow the same steps as the previous updates.

Inside the `Naive.Trader` module we need to update all the references to the `Logger` to an `@logger` attribute with value dependent on the configuration:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/trader.ex
  ...
  @logger Application.compile_env(:core, :logger)
  ...
    @logger.info("Initializing new trader(#{id}) for #{symbol}")
  ...
    @logger.info(
      "The trader(#{id}) is placing a BUY order " <>
        "for #{symbol} @ #{price}, quantity: #{quantity}"
    )
  ...
        @logger.info(
          "The trader(#{id}) is placing a SELL order for " <>
            "#{symbol} @ #{sell_price}, quantity: #{quantity}."
        )
  ...
      @logger.info("Trader(#{id}) finished trade cycle for #{symbol}")
  ...
      @logger.info("Rebuy triggered for #{symbol} by the trader(#{id})")
  ...
```

The `Logger` module doesn't provide a behaviour. As we can't modify its source,
we need to create a new module to define the Logger behaviour.
Let's place it inside the `Core.Test` namespace in the `/apps/core/lib/core/test.ex` file
side by side with the PubSub behaviour with the following definition:

```{r, engine = 'elixir', eval = FALSE}
# /apps/core/lib/core/test.ex
defmodule Core.Test do
  ...
  defmodule Logger do
    @type message :: binary

    @callback info(message) :: :ok
  end
end
```

As previously, we defined a callback and additional type for the `message` argument.

Next, we will use the above behaviour to define a mock using the `Mox.defmock/2` function inside the `test_helper.exs` file of the `naive` application:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/test_helper.exs
Mox.defmock(Test.LoggerMock, for: Core.Test.Logger)
``` 

In the above code, we've told the `mox` package to define the `Test.LoggerMock` module based on the `Core.Test.Logger` behaviour.



The final step will be to append the `:core, :logger` configuration to the `/config/config.exs` file:

```{r, engine = 'elixir', eval = FALSE}
# /config/config.exs
config :core,                                   
  logger: Logger,                # <= added
  pubsub_client: Phoenix.PubSub
``` 

and the test `/config/test.exs` configuration file:

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
config :core,                                   
  logger: Test.LoggerMock,        # <= added
  pubsub_client: Test.PubSubMock
```

This finishes mocking all four dependencies.
Every external module the `Naive.Trader` relies on is now based on configuration values.
We can progress to writing the test.

## Implement a test of the `Naive.Trader` module 

Finally, we can implement the unit test for the `Naive.Trader` module.

We will create a folder called `naive` inside the `/apps/naive/test` directory and a new file called `trader_test.exs` inside it. 

Let's start with an empty skeleton of the test tagged as `unit`:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive/trader_test.exs
defmodule Naive.TraderTest do
  use ExUnit.Case
  doctest Naive.Trader

  @tag :unit
  test "Placing buy order test" do
  end
end
```

Let's add the `mox` related code above the `@tag :unit` line:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive/trader_test.exs
  import Mox                   # <= 1 

  setup :set_mox_from_context  # <= 2
  setup :verify_on_exit!       # <= 3
```

In the above code, we are:

- importing the `mox` module so we will be able to use functions like `expect/3`
- we use `set_mox_from_context` because our test is not running in `async: true mode`, Mox allows us to share expectations globally. This is crucial because the `Naive.Trader` runs in a separate process from the test. If we ran this test asynchronously, the Trader process would crash because it wouldn't have permission to call the mocks defined by the test process(unless we explicitly granted allowances).
- telling `mox` to verify that all the mocks defined in the tests have been called from within those tests. Otherwise, it will flag such cases as test errors

Inside our test, we need to define implementations for all the functions that the `Naive.Trader` relies on:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive/trader_test.exs
  ...
  test "Placing buy order test" do
    Test.PubSubMock
    |> expect(:subscribe, fn _module, "TRADE_EVENTS:XRPUSDT" -> :ok end) # <= 1
    |> expect(:broadcast, fn _module, "ORDERS:XRPUSDT", _order -> :ok end)

    Test.BinanceMock
    |> expect(:order_limit_buy, fn "XRPUSDT", "464.360", 0.4307, "GTC" -> # <= 2
      {:ok,
       BinanceMock.generate_fake_order(
         "XRPUSDT",
         "464.360",
         0.4307,
         "BUY"
       )
       |> BinanceMock.convert_order_to_order_response()}
    end)

    test_pid = self() # <= 3

    Test.Naive.LeaderMock
    |> expect(:notify, fn :trader_state_updated, %Naive.Trader.State{} ->
      send(test_pid, :ok) # <= 3
      :ok
    end)

    Test.LoggerMock
    |> expect(:info, 2, fn _message -> :ok end) # <= 4
    ...
```

It's important to note that we defined the mocked function with expected values in the above code.
We expect our test to subscribe to the trade events topic and broadcast to the orders topic(#1).
We are also expecting that the process will place an order at the exact values that we calculated upfront.
This way, our mock becomes an integral part of the test,
asserting that the correct values will be passed to other parts of the system(dependencies of the `Naive.Trader` module).

Another "trick"(#3) that we can use in our mocks is to leverage the fact that we can send
a message to the test process from within the mocked function.
We will use this idea to know precisely when the trader process finished its work as the `notify/2` is
the last function call inside the process' callback(`handle_info/2` inside the `Naive.Trader` module).
We will assert that we should receive the message, and the test will be waiting for it before exiting(the default timeout is 100ms)
instead of using hardcoded `sleep` to "hack" it to work.

The final part(#4) tells the `mox` package that `Logger.info/1` will be called twice inside the test.
The `mox` will verify the number of calls to the mocked function and error if it doesn't match the expected amount.

The second part of the test is preparing the initial state for the `Naive.Trader` process, generating a trade event and sending it to the process:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive/trader_test.exs
  ...
  test "Placing buy order test" do
    ...
    trader_state = dummy_trader_state()
    trade_event = generate_event(1, 0.4318301, "213.10000000")

    {:ok, trader_pid} = Naive.Trader.start_link(trader_state)
    send(trader_pid, trade_event)
    assert_receive :ok
  end
```

As described above, the `assert_receive/1` function will cause the test to wait for the message for 100ms before quitting.

Here are the helper functions that we used to generate the initial trader state and trade event:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive/trader_test.exs
  ...
  test "Placing buy order test" do
    ...
  end

  defp dummy_trader_state() do
    %Naive.Trader.State{
      id: 100_000_000,
      symbol: "XRPUSDT",
      budget: "200",
      buy_order: nil,
      sell_order: nil,
      buy_down_interval: Decimal.new("0.0025"),
      profit_target: Decimal.new("0.001"),
      rebuy_interval: Decimal.new("0.006"),
      rebuy_notified: false,
      tick_size: "0.0001",
      step_size: "0.001"
    }
  end

  defp generate_event(id, price, quantity) do
    %Core.Struct.TradeEvent{
      event_type: "trade",
      event_time: 1_000 + id * 10,
      symbol: "XRPUSDT",
      trade_id: 2_000 + id * 10,
      price: price,
      quantity: quantity,
      trade_time: 5_000 + id * 10,
      buyer_market_maker: false
    }
  end
```

The above code finishes the implementation of the test, but inside the test we used functions from the `BinanceMock` module that are currently private.

We need to update the module by making the `generate_fake_order/4` and
`convert_order_to_order_response/1` functions public(and moving them up in the module, so they are next to other public functions):

```{r, engine = 'elixir', eval = FALSE}
# /apps/binance_mock/lib/binance_mock.ex
  ...
  def get_order(symbol, time, order_id) do
    ...
  end

  def generate_fake_order(symbol, quantity, price, side) # <= updated to public
    ...
  end

  def convert_order_to_order_response(%Binance.Order{} = order) do # <= updated to public
    ...
  end
...
```

We updated both of the methods to public and moved them up after the `get_order/3` function.

## Define an alias to run unit tests

Our unit test should be run without running the whole application, so we need to run it with the `--no-start` argument.
We should also select unit tests by tag(`--only unit`). Let's create an alias that will hide those details:

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  defp aliases do
    [
      ...
      "test.unit": [
        "test --only unit --no-start"
      ]
    ]
  end
```

We can now run our test using a terminal:

```{r, engine = 'bash', eval = FALSE}
MIX_ENV=test mix test.unit
```

We should see the following error:

```{r, engine = 'bash', eval = FALSE}
21:22:03.811 [error] GenServer #PID<0.641.0> terminating
** (stop) exited in: GenServer.call(BinanceMock, :generate_id, 5000)
    ** (EXIT) no process: the process is not alive or there's no process currently
       associated with the given name, possibly because its application isn't started
```

One of the `BinanceMock` module's functions is sending a message to generate a unique id to the process
that doesn't exist(as we are running our tests without starting the supervision tree[the `--no-start` argument]).

There are two ways to handle this issue:

- inside the `/apps/naive/test/test_helper.exs` file we could ensure that the `BinanceMock` is up and running by adding `Application.ensure_all_started(:binance_mock)` function call - this is a hack
- we could refactor the `BinanceMock.generate_fake_order/4` to accept `order_id` as an argument instead of sending the message internally - this should be a much better solution. Let's give it a shot.

First, let's update the `BinanceMock` module:

```{r, engine = 'elixir', eval = FALSE}
# /apps/binance_mock/lib/binance_mock.ex
  def generate_fake_order(order_id, symbol, quantity, price, side) # <= order_id added
    ...
    # remove the call to GenServer from the body
    ...
  end
  ...
  defp order_limit(symbol, quantity, price, side) do
    ...
      generate_fake_order(
        GenServer.call(__MODULE__, :generate_id), # <= order_id generation added
        symbol,
        quantity,
        price,
        side
      )
```

Now we need to update our test to pass some dummy order id from the mocked function:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive/trader_test.exs
  ...
  test "Placing buy order test" do
    ...
      {:ok, BinanceMock.generate_fake_order(
        "12345",                        # <= order_id added
        "XRPUSDT",
        "464.360",
        0.4307,
        "BUY"
      )
   ...
  end
```

We can now rerun our test:

```{r, engine = 'bash', eval = FALSE}
MIX_ENV=test mix test.unit
...
Finished in 0.1 seconds (0.00s async, 0.1s sync)
1 tests, 0 failures (1 excluded)
```

Congrats! We just successfully tested placing an order without any dependencies.

To avoid explicitly passing the `MIX_ENV=test` environment variable, we will add the preferred environment for our alias inside the `mix.exs` file:

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  # add below under the `project` function
  def cli do
    [preferred_envs: ["test.unit": :test]]
  end
```

Now we can run our tests by:

```{r, engine = 'bash', eval = FALSE}
mix test.unit
...
Finished in 0.06 seconds (0.00s async, 0.06s sync)
1 tests, 0 failures (1 excluded)
```

**We now have fast, isolated unit tests powered by `mox`.** Here's what we achieved:

- Defined behaviours for modules that lacked them (`BinanceMock`, `Naive.Leader`, `Phoenix.PubSub`, `Logger`)
- Used `Mox.defmock/2` to generate mock modules constrained by those behaviours
- Wired up environment-based configuration so tests use mocks while dev uses real modules
- Built a test for `Naive.Trader` that verifies the exact values passed to every dependency
- Eliminated hardcoded `sleep` calls by using a message-passing feedback loop from mock to test

The key insight is that `mox` turns your mocks into assertions.
Instead of just stubbing out dependencies, each mock expectation verifies
that the right function was called with the right arguments.
Your mocks are doing double duty - enabling isolation *and* validating correctness.

It's worth noting the trade-offs of the `mox` approach:

- behaviours must be defined for every module you want to mock - this is upfront work, but it pays off as a form of documentation
- module attributes are resolved at compile time, so you can't swap implementations mid-test without recompilation
- mocked tests don't catch integration issues - that's what our end-to-end test from the last chapter is for

The two testing approaches complement each other: end-to-end tests prove the system works,
`mox` tests prove each piece works correctly in isolation.

**So, What's Next?**

We've been building OTP processes and testing them, but we haven't talked much about
functional programming itself. In the next chapter, we'll look at how to separate
pure business logic from side effects.
We'll extract our trading strategy's decision-making into pure functions
that are trivial to test - no mocks needed at all.

[Note] Please remember to run `mix format` to keep things nice and tidy.

The source code for this chapter can be found in the book's source code repository
(branch:
[chapter_16](https://github.com/Cinderella-Man/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_16)).
