# Functional Elixir

In the last chapter, we built fast, isolated unit tests using `mox`. 
We mocked dependencies, verified interactions, and eliminated our reliance 
on databases and sleeps. The tests run in milliseconds - but here's the uncomfortable truth: 
we're still testing a lot of side effects.

Every expectation we define with `mox` is essentially saying "this function 
should call that external system with these exact values." We've isolated our code, 
but we haven't fundamentally changed what it does - it's still all side effects, 
all the way down.

What if we could write code that doesn't need mocking at all? Code so pure that 
testing it is trivial - just call the function and check the output. No setup, 
no mocks, no expectations. Just input → function → output.

That's what functional programming gives us. We'll extract the decision-making 
logic from our `Naive.Trader` into pure functions that take state and events, 
return decisions. No PubSub calls, no Binance API, no database - just math.

Along the way, we'll see why Elixir's approach to functional programming is 
pragmatic rather than dogmatic. We'll use `with` for error handling, module 
attributes for dependency injection, and we'll draw a hard line between 
"functional" as a useful practice and "functional" as academic category theory.

## Objectives
- the reasoning behind the functional approach
- simplifying by splitting
- abstracting the "pure" logic
- dealing with dirty code
- making dirty code testable
- the power `with`-in
- to do or not to do - side effects at the edge
- final thoughts

## The reasoning behind the functional approach

Across the last 16 chapters, we focused on learning OTP by building a trading system.
We deliberately delayed the conversation about functional programming to maintain a clear focus on OTP.
But here's the thing - we've been writing functional code all along.

We use higher-order functions, pattern matching, immutable data. Isn't that functional enough?

Not quite. We're using functional *syntax*, but we haven't embraced functional *design*. 
Our functions mix business logic with side effects. They make decisions and execute them in the same breath. 
This makes them harder to test and harder to reason about.

The real value of functional programming isn't academic purity - it's practical testability. 
Pure functions are trivial to test: call them with inputs, verify outputs. No mocks, no setup, no cleanup. 

We'll start from the basics and examine different ways to implement functional concepts, 
always considering Elixir's specific strengths and weaknesses.

## Simplifying by splitting

Note: This refactoring might seem tangential to functional programming, but we're doing it now to avoid 
complexity later. When we extract pure logic in the next section, these simplified callbacks will be 
much easier to work with.

Let's look at our strategy inside the `Naive.Trader` module.
In this section, we will focus on its `handle_info/2` callback function.

We are looking for clauses that do more than "one thing" to split them into multiple clauses:

- The first callback places a buy order - it has a single responsibility and is easy to follow.
- The second callback is the one we will focus on. It deals with two things, and we could describe it as a "fetch buy order and place a sell order" function. The "and" in the description clearly indicates that it's really two functions glued together. We will split it into "fetch buy order" and "place a sell order" functions(below code replaces the 2nd `handle_info/2` callback):

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def handle_info(
        %TradeEvent{},
        %State{
          id: id,
          symbol: symbol,
          buy_order: %Binance.OrderResponse{
            price: buy_price,
            orig_qty: quantity,
            status: "FILLED"
          },
          sell_order: nil,
          profit_target: profit_target,
          tick_size: tick_size
        } = state
      ) do
    sell_price = calculate_sell_price(buy_price, profit_target, tick_size)

    @logger.info(
      "The trader(#{id}) is placing a SELL order for " <>
        "#{symbol} @ #{sell_price}, quantity: #{quantity}."
    )

    {:ok, %Binance.OrderResponse{} = order} =
      @binance_client.order_limit_sell(symbol, quantity, sell_price, "GTC")

    :ok = broadcast_order(order)

    new_state = %{state | sell_order: order}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end

  def handle_info(
        %TradeEvent{
          price: trade_price
        },
        %State{
          id: id,
          symbol: symbol,
          buy_order: %Binance.OrderResponse{
            price: buy_price,
            order_id: order_id,
            transact_time: timestamp
          },
          sell_order: nil
        } = state
      )
      when trade_price <= buy_price do
    @logger.info("Trader's(#{id}) #{symbol} buy order got filled")

    {:ok, %Binance.Order{} = current_buy_order} =
      @binance_client.get_order(
        symbol,
        timestamp,
        order_id
      )

    buy_order_response = convert_order_to_order_response(current_buy_order)
    :ok = broadcast_order(buy_order_response)

    new_state = %{state | buy_order: buy_order_response}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end
```

The first function takes care of placing a sell order. The second one fetches the buy order -
please note that we needed to enhance the condition to avoid the 

We can now move to the next clause.
Similar to the last one, we could describe it as "fetch the sell order and terminate the trader".
We will split it into two callbacks: "fetch the sell order" and "terminate trader".

\newpage

The code below replaces the 5th `handle_info/2` callback:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def handle_info(
        %TradeEvent{},
        %State{
          id: id,
          symbol: symbol,
          sell_order: %Binance.OrderResponse{
            status: "FILLED"
          }
        } = state
      ) do
    @logger.info("Trader(#{id}) finished trade cycle for #{symbol}")
    {:stop, :normal, state}
  end

  def handle_info(
        %TradeEvent{
          price: trade_price
        },
        %State{
          id: id,
          symbol: symbol,
          sell_order: %Binance.OrderResponse{
            price: sell_price,
            order_id: order_id,
            transact_time: timestamp
          }
        } = state
      )
      when trade_price >= sell_price do
    @logger.info("Trader's(#{id}) #{symbol} SELL order got filled")

    {:ok, %Binance.Order{} = current_sell_order} =
      @binance_client.get_order(
        symbol,
        timestamp,
        order_id
      )

    sell_order_response = convert_order_to_order_response(current_sell_order)
    
    :ok = broadcast_order(sell_order_response)

    new_state = %{state | sell_order: sell_order_response}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end
```

The first function takes care of terminating the trader. The second function is fetching the sell order.

That finishes our first refactoring round, but I need to admit that our change has impacted the behaviour of our strategy.
Each time a buy or sell order gets filled, we will fetch that order from Binance,
but we **won't** immediately place a sell order nor terminate as it was happening before.
Instead, only when another event arrives will the trader place a sell order or terminate.

Changes like this require approval from the business in a work situation,
but it's a good showcase of the situation where we can propose a solution
that will simplify the code(the benefits will become evident in the following sections).

We can confirm that we have broken our tests by running our integration testsuite:

```{r, engine = 'bash', eval = FALSE}
$ MIX_ENV=integration mix test.integration
...
  1) test Naive trader full trade(buy + sell) test (NaiveTest)
     apps/naive/test/naive_test.exs:12
     Assertion with == failed
     code:  assert buy_2 == [Decimal.new("0.4309"), "BUY", "NEW"]
     left:  [Decimal.new("0.431"), "BUY", "NEW"]
     right: [Decimal.new("0.4309"), "BUY", "NEW"]
     stacktrace:
       test/naive_test.exs:85: (test)
```

(Subject to acceptance by the business) we will fix the integration test in the following way:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/test/naive_test.exs
  test "Naive trader full trade(buy + sell) test" do
    ...
    # events' comments updated below
    # Step 4 - Broadcast 8 events

    [
      # below event will trigger
      # buy order placed @ 0.4307
      generate_event(1, 0.43183010, "213.10000000"),
      # above the buy price - ignored
      generate_event(2, 0.43183020, "56.10000000"),
      # above the buy price - ignored
      generate_event(3, 0.43183030, "12.10000000"),
      # event at the expected buy price
      # it should trigger fetching the buy order
      generate_event(4, 0.4307, "38.92000000"),
      # event below the expected buy price
      # normally ignored but after fetching the buy order
      # it should trigger placing a sell order @ 0.4319
      generate_event(5, 0.43065, "126.53000000"),
      # event at exact the expected sell price
      # it should trigger fetching the sell order
      generate_event(6, 0.4319, "62.92640000"),
      # event after fetching the sell order
      # causes trader process to exit
      generate_event(7, 0.43205, "345.14235000"),
      # below event will trigger
      # buy order placed @ 0.431
      generate_event(8, 0.43210, "3201.86480000")
    ]
    ...
    assert buy_2 == [Decimal.new("0.431"), "BUY", "NEW"] # <- updated from 0.4309
```

We have updated the 2nd order buy price to make our test green again.

\newpage

## Abstracting the "pure" logic

In our adventure to make our code more functional, we should strive
to separate(as much as possible) pure business logic from side effects and boilerplate.

The `Naive.Trader` module is a GenServer that receives trade events via messages.
Based on them and the current state, using pattern-matching, it decides what action should be performed(place a buy order,
fetch a buy order, place a sell order, fetch sell order, terminate trader, trigger rebuy or ignore event).

Each of the pattern-matches inside the callback functions' headers is a **strategy** specific business logic
that got mixed with the fact that it's executed by a GenServer that receives messages.

We will create a new file called `strategy.ex` inside the `apps/naive/lib/naive/` directory,
where we will **copy** all of the `handle_info/2` callback functions from the `Naive.Trader` module:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
defmodule Naive.Strategy do

  def handle_info(...) do
    ... # <= place a buy order logic
  end

  def handle_info(...) do
    ... # <= place a sell order logic
  end

  def handle_info(...) do
    ... # <= fetch the buy order logic
  end

  def handle_info(...) do
    ... # <= terminate trader logic
  end

  def handle_info(...) do
    ... # <= fetch the sell order logic
  end

  def handle_info(...) do
    ... # <= trigger rebuy order logic
  end

  def handle_info(...) do
    ... # <= ignore trade event logic
  end
```

First, we will rename all of the `handle_info/2` functions inside the `Naive.Strategy` module to `generate_decision/2`.
Next, we will go through them one by one, leaving the pure parts and limiting them to returning the decision.

### Place a buy order rules

The first function decides should the trader place a buy order. We can see that price
and quantity calculations are pure functions based on the incoming data. We will remove everything below those two as it's causing side effects.

As now we are dealing with a function generating a decision, we will return a tuple with data that,
together with state, will be used to place a buy order. 

After removing some of the pattern-matching that we used to retrieve data(no longer needed),
our first function should look like this:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the first clause
  def generate_decision(
        %TradeEvent{price: price},
        %State{
          budget: budget,
          buy_order: nil,
          buy_down_interval: buy_down_interval,
          tick_size: tick_size,
          step_size: step_size
        }
      ) do
    price = calculate_buy_price(price, buy_down_interval, tick_size)
    quantity = calculate_quantity(budget, price, step_size)

    {:place_buy_order, price, quantity}
  end
```
### Place a sell order rules

We will follow the same logic for the 2nd clause of the `generate_decision/2` function.
We will leave only the sell price calculation as it's pure and return a tuple together with the decision:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the second clause
  def generate_decision(
        %TradeEvent{},
        %State{
          buy_order: %Binance.OrderResponse{
            price: buy_price,
            status: "FILLED"
          },
          sell_order: nil,
          profit_target: profit_target,
          tick_size: tick_size
        }
      ) do
    sell_price = calculate_sell_price(buy_price, profit_target, tick_size)
    {:place_sell_order, sell_price}
  end
```

### Fetch the buy order rules

For the 3rd clause, we will return only an atom as there's no pure logic besides the pattern-match in the header itself:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the third clause
  def generate_decision(
        %TradeEvent{
          price: trade_price
        },
        %State{
          buy_order: %Binance.OrderResponse{
            price: buy_price
          },
          sell_order: nil
        }
      )
      when trade_price <= buy_price do
    :fetch_buy_order
  end
```

### Terminate trader rules

For the 4th clause, we will indicate that trader needs to terminate:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the fourth clause
  def generate_decision(
        %TradeEvent{},
        %State{
          sell_order: %Binance.OrderResponse{
            status: "FILLED"
          }
        }
      ) do
    :exit
  end
```

### Fetch the sell order rules

For the 5th clause, we will indicate that trader needs to fetch the sell order:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the fifth clause
  def generate_decision(
        %TradeEvent{
          price: trade_price
        },
        %State{
          sell_order: %Binance.OrderResponse{
            price: sell_price
          }
        }
      )
      when trade_price >= sell_price do
    :fetch_sell_order
  end
```

### Trigger rebuy rules

Inside the 6th clause, we are dealing with triggering the rebuy.
Here, we can decide whether rebuy should be triggered and get rid of conditional logic inside further steps.
We couldn't refactor this function by splitting it (as we've done in the first section)
as we need to call the `trigger_rebuy?/3` function to check should rebuy be triggered.
The functions that we refactored in the first section of this chapter were splittable
as they relied on pattern-matching in the function headers where calling local functions is not allowed):

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the sixth clause
  def generate_decision(
        %TradeEvent{
          price: current_price
        },
        %State{
          buy_order: %Binance.OrderResponse{
            price: buy_price
          },
          rebuy_interval: rebuy_interval,
          rebuy_notified: false
        }
      ) do
    if trigger_rebuy?(buy_price, current_price, rebuy_interval) do
      :rebuy
    else
      :skip
    end
  end
```

### The final clause rules

The final (7th) clause will just ignore the trade event as it's of no interest:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# the final(7th) clause
  def generate_decision(%TradeEvent{}, _state) do
    :skip
  end
```

This finishes the changes to the `generate_decision/2` clauses.
We extracted a fair amount of logic into an easily testable pure function. We now need to use it inside the `Naive.Trader` module.

### Changes to the `Naive.Trader` module

We will start by moving all of the calculation functions to the `Naive.Strategy` module as
we are using them from the `generate_decision/2` function. Those will be:

- `calculate_sell_price/3`
- `calculate_buy_price/3`
- `calculate_quantity/3`
- `trigger_rebuy?/3`

They can now be changed to public functions as they are pure and fit the "interface"
of the `Naive.Strategy` module(it feels ok[and it's safe as they are pure] to "expose" them to be called from other modules).

We need to remember about moving the `Decimal` alias into the `Naive.Strategy` module together with
a copy of the `TradeEvent` struct alias and add the alias for the `Naive.Trader.State` struct:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
  alias Decimal, as: D
  alias Core.Struct.TradeEvent
  alias Naive.Trader.State
```

The next step will be to rename all the `handle_info/2` callback functions inside the `Naive.Trader` module
to `execute_decision/2`, which we will get back to in a moment.

First, we need to add a single `handle_info/2` callback under the `init/1` function that will pattern match
only the fact that the received message contains the `TradeEvent` struct and the state is the correct `State` struct:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
# add after the `init/1` function
  def handle_info(%TradeEvent{} = trade_event, %State{} = state) do
    Naive.Strategy.generate_decision(trade_event, state)
    |> execute_decision(state)
  end
```

So, the `Naive.Strategy` module will decide what the trader server should do based on its pure business logic.
That decision will be passed forward with the state to the `execute_decision/2` function (at this moment,
it's just the old `handle_info/2` function renamed, but we will update it next).

### `Naive.Trader` - place buy order

We will update the `execute_decision/2` function to take a decision + state and execute the correct
action based on pattern-match of the decision. Starting with the 1st clause, we need to pattern match a tuple:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
# the first execute clause
  def execute_decision(
        {:place_buy_order, price, quantity},
        %State{
          id: id,
          symbol: symbol
        } = state
      ) do
    @logger.info(
      "The trader(#{id}) is placing a BUY order " <>
        "for #{symbol} @ #{price}, quantity: #{quantity}"
    )

    {:ok, %Binance.OrderResponse{} = order} =
      @binance_client.order_limit_buy(symbol, quantity, price, "GTC")

    :ok = broadcast_order(order)

    new_state = %{state | buy_order: order}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end
```

The amount of pattern matching will be much smaller as part of the original callback has been moved inside
the `Naive.Strategy`'s logic(to calculate the price and quantity).

### `Naive.Trader` - Place a sell order

In case of placing a sell order, we will pattern match on a tuple containing the `:place_sell_order` atom and slim down on pattern matching:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def execute_decision(
        {:place_sell_order, sell_price},
        %State{
          id: id,
          symbol: symbol,
          buy_order: %Binance.OrderResponse{
            orig_qty: quantity
          }
        } = state
      ) do
    @logger.info(
      "The trader(#{id}) is placing a SELL order for " <>
        "#{symbol} @ #{sell_price}, quantity: #{quantity}."
    )

    {:ok, %Binance.OrderResponse{} = order} =
      @binance_client.order_limit_sell(symbol, quantity, sell_price, "GTC")

    :ok = broadcast_order(order)
    new_state = %{state | sell_order: order}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end
```

### `Naive.Trader` - Fetch the buy order

In case of fetching the buy order, we will pattern match on a `:fetch_buy_order` atom and slim down on pattern matching:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def execute_decision(
        :fetch_buy_order,
        %State{
          id: id,
          symbol: symbol,
          buy_order: %Binance.OrderResponse{
            order_id: order_id,
            transact_time: timestamp
          }
        } = state
      ) do
    @logger.info("Trader's(#{id}) #{symbol} buy order got filled")

    {:ok, %Binance.Order{} = current_buy_order} =
      @binance_client.get_order(
        symbol,
        timestamp,
        order_id
      )

    buy_order_response = convert_order_to_order_response(current_buy_order)
    :ok = broadcast_order(buy_order_response)
    new_state = %{state | buy_order: buy_order_response}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end
```

### `Naive.Trader` - Terminate the trader

In case of terminating the trader, we will pattern match on a `:exit` atom and slim down on pattern matching:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def execute_decision(
        :exit,
        %State{
          id: id,
          symbol: symbol
        } = state
      ) do
    @logger.info("Trader(#{id}) finished trade cycle for #{symbol}")
    {:stop, :normal, state}
  end
```

\newpage

### `Naive.Trader` - Fetch the sell order

In case of fetching the sell order, we will pattern match on a `:fetch_sell_order` atom and slim down on pattern matching:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def execute_decision(
        :fetch_sell_order,
        %State{
          id: id,
          symbol: symbol,
          sell_order: %Binance.OrderResponse{
            order_id: order_id,
            transact_time: timestamp
          }
        } = state
      ) do
    @logger.info("Trader's(#{id}) #{symbol} SELL order got filled")

    {:ok, %Binance.Order{} = current_sell_order} =
      @binance_client.get_order(
        symbol,
        timestamp,
        order_id
      )

    sell_order_response = convert_order_to_order_response(current_sell_order)
    :ok = broadcast_order(sell_order_response)
    new_state = %{state | sell_order: sell_order_response}
    @leader.notify(:trader_state_updated, new_state)
    {:noreply, new_state}
  end
```

### `Naive.Trader` - Triggering rebuy

In case of triggering the rebuy procedure, we will pattern match on a `:rebuy` atom, slim down on pattern matching and
simplify the function a fair bit(no branching required anymore - yay!):

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def execute_decision(
        :rebuy,
        %State{
          id: id,
          symbol: symbol
        } = state
      ) do
    @logger.info("Rebuy triggered for #{symbol} by the trader(#{id})")
    new_state = %{state | rebuy_notified: true}
    @leader.notify(:rebuy_triggered, new_state)
    {:noreply, new_state}
  end
```

### `Naive.Trader` - The final ignore clause

The final ignore clause will `:skip` all events:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def execute_decision(:skip, state) do
    {:noreply, state}
  end
```

The ignore clause finishes our current refactoring round, which showcased that sometimes abstracting pattern matching into
a separate function is a valid strategy to increase the amount of pure code.

Note: The fact that we could abstract the logic from pattern matches is quite a unique situation to our application.
I would not advise abstracting GenServer pattern matching into a separate module if dealing with different structs/actions(in our case,
all our pattern matches were "making a trading decision", it's a single "action", that's why we abstracted them).

## Dealing with dirty code

In the last section, we've split the `handle_info/2` clauses into the `generate_decision/2` and `execute_decision/2` functions.
That's excellent progress, but we still have the strategy logic inside the `Naive.Trader` module.

Let's move the `execute_decision/2` function(together with all the code that it depends on, like the `convert_order_to_order_response/1`,
`broadcast_order/1`  and `convert_to_order/1` functions as well as a copy of the `require Logger`) from the `Naive.Trader` module
to the `Naive.Strategy` module.

As the `execute_decision/2` function is causing side effects, we don't want it to be called directly from the outside of the module,
so we will need to make it private.

Changing the `execute_decision/2` function(now inside the `Naive.Strategy` module) to private will cause
a problem with the `handle_info/2` callback function inside the `Naive.Trader` module as
it relies on the `execute_decision/2` function to be public.
The fact that our strategy makes a decision and then executes code based on it is an implementation detail that
we shouldn't share with the `Naive.Trader` module. That's why we will move the **body** of the `handle_info/2` callback function
into a new function called `execute/2` inside the `Naive.Strategy` module:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
  def execute(%TradeEvent{} = trade_event, %State{} = state) do
    generate_decision(trade_event, state)
    |> execute_decision(state)
  end
```

Before updating the `Naive.Trader` module to use the `execute/2` function, we need to address another issue that
moving the `execute_decision/2` caused. At this moment, all of the clauses return GenServer specific tuples.
What we really need to return to the trader is an atom indicating should it continue or terminate together with the updated state:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# last lines inside the `execute_decision/2` clauses
{:ok, new_state} # <= previously {:noreply, new_state} (5 times)
{:ok, state} # <= previously {:noreply, state} (once)
:exit # <= previously {:stop, :normal, state} + remove `state` pattern match (once)
```

We can now update the `handle_info/2` callback function to call the new "interface" of the `Naive.Strategy` module
that we just created and act accordingly to the result:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/trader.ex
  def handle_info(%TradeEvent{} = trade_event, %State{} = state) do
    case Naive.Strategy.execute(trade_event, state) do
      {:ok, new_state} -> {:noreply, new_state}
      :exit -> {:stop, :normal, state}
    end
  end
```

At this moment, we could just copy/move module attributes from the `Naive.Trader` module to the `Naive.Strategy` module
and our code would start to work again. Still, before we will do that, we will use this opportunity to look into how to make our dirty code testable.

## Making dirty code testable

Making dirty code testable is very closely linked to injecting dependencies. In the testing environment,
we would like to use dummy implementations instead of executing the side-effect-causing code to simplify the tests.
We will look into the different ways that we can pass side-effect-causing "code" into impure functions.

### Passing functions  arguments

Functions are first-class citizens in Elixir, which means that we can pass them as arguments to functions.
This way, we can pass side-effect causing functions into our `Naive.Strategy` module.

Let's look at how this would look in practice. We need to look into the `execute_decision/2` function,
as it's the place where side effects happen. Looking at the 1st clause(responsible for placing a buy order),
we can see that it's calling the `Logger.info/1`, `Binance.order_limit_buy/4`, `PubSub.broadcast/3`(via the `broadcast_order/1` function)
and `Leader.notify/2` functions. To make our code easily testable, we would need to be able to pass dummy implementations for all of those.

As we aren't calling the `execute_decision/2` directly, we need to pass all of the above functions as arguments
to the `execute/2` function, which will pass them onward to the `execute_decision/2`.

We can see that even with default values pointing to the "real" implementation, that's still **a lot** of noise to make testing easier.
It will negatively impact the maintenance of the code - here's an example of what this would look like(don't bother typing it):

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
# injecting dummy implementation, fallback to real implementation
def execute(
  %TradeEvent{} = trade_event,
  %State{} = state,
  logger_info \\ &Logger.info/1, # <= function injected
  order_limit_buy \\ &Binance.order_limit_buy/4, # <= function injected
  pubsub_broadcast \\ &PubSub.broadcast/3, # <= function injected
  notify_leader \\ &Leader.notify/2 # <= function injected
) do
  ...
```

There are already four functions, and we only took care of side-effects causing functions from the first `execute_decision/2` clause.
We can easily see how this very quickly becomes just unmanageable as there would be 10+ "injected"
arguments going from the `execute/2` to `execute_decision/2`, and only some of them would be used in each clause.

Additional downsides:

- when passing a function as an argument, we need to specify the arity, so when we would like to use more than one arity, we need to pass the function **multiple times with different arities**. An example could be passing `Logger.info/1` and `Logger.info/2`
- we need to give a name to every passed function, sometimes multiple arities (again, how should variables for `Logger.info/1` and `Logger.info/2` be called? `logger_info_2`?)
- share amount of arguments negatively impacts code readability

We can see that passing functions as arguments is just a bad idea in case of making our code testable.
It will have the opposite effect, decreasing readability making our code difficult to maintain and follow.

Important note: Passing functions as arguments is not always bad! A good example could be when different actions need to be performed based on runtime data.

\newpage

### Passing grouped functions as a context

The natural next step would be to put all of those functions into some structure like Map or Keyword list.
Whichever we would choose, we will end up with the same problems of naming keys(this time inside the map/keyword list),
multiple functions because of different arity but also default values inside each clause of the `execute_decision/2` function:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
  defp execute_decision(
        {:place_buy_order, price, quantity},
        %State{
          id: id,
          symbol: symbol
        } = state,
        %{} = context # <= context added
      ) do
    # vvv fetch from context vvv
    logger_info = Map.get(context, :logger_info, &Logger.info/1)
    order_limit_buy = Map.get(context, :order_limit_buy, &Binance.order_limit_buy/4)
    leader_notify = Map.get(context, :leader_notify, &Leader.notify/2)
```

Again this looks like a bad idea. It's probably marginally better than just sending functions one by one, but not much.

### Passing grouped modules as a context

The significant advantage of passing modules as arguments instead of functions is that we no longer
have a problem with naming keys or caring about different functions' arities.
There will also be substantially fewer modules used in comparison to functions.

Sadly we still need to use the `Map` function to get the modules out of `"context"`:

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
  defp execute_decision(
        {:place_buy_order, price, quantity},
        %State{
          id: id,
          symbol: symbol
        } = state,
        %{} = context # <= context added
      ) do
    logger = Map.get(context, :logger, &Logger) # <= fetch from context
    binance = Map.get(context, :binance, Binance) # <= fetch from context
    leader = Map.get(context, :leader, &Leader) # <= fetch from context
```

This is much better, but we will still need to do a fair amount of additional work to get the modules out.
Also, our code will be full of the "default" modules(as **each** of the clauses retrieving them from the context will need to specify defaults).

### Injecting modules to module's attributes based on the configuration

And we finally got there - we have come a full circle.
This is the approach that we previously used inside the `Naive.Trader` module(you can go ahead and
add them to the `Naive.Strategy` module as well as remove the `@binance_client` and `@leader`
from the `Naive.Trader` module as they are no longer needed):

```{r, engine = 'elixir', eval = FALSE}
# /apps/naive/lib/naive/strategy.ex
defmodule Naive.Strategy do
  ...

  @binance_client Application.compile_env(:naive, :binance_client)
  @leader Application.compile_env(:naive, :leader)
  @logger Application.compile_env(:core, :logger)
  @pubsub_client Application.compile_env(:core, :pubsub_client)
```

We looked into different ways to inject dependencies to understand their downsides.

Sometimes, injecting values(like modules) as module attributes can feel like a "global state", "singleton", or similar antipattern.
We need to understand that each programming language provides different ways to solve common programming problems.
Dependency injection is one of those common concerns that every language needs to solve, and Elixir solves it by using compile-time modules' attributes.

As long as you use module attributes to inject compile-time dependencies for testing,
there's just **no better way** to do it in Elixir, and now we know why(based on the issues with the alternative approaches).

Together with the module attributes, our code should be now fully functional.

## The power `with`-in

In the last section, we looked into different ways to inject modules' dependencies to avoid side-effect-causing functions inside the tests.
Besides side-effect causing functions, in functional programming, error handling is also done in a specific manner.

Many languages introduced concepts like `Either`, which is a struct that can be either `Left`(error result) or `Right`(success result).
Those quite nicely fit to the standard Elixir results like `{:error, reason}` and `{:ok, result}`.
Further, those languages provide multiple functions to work with the `Either`, like `map`.

```{r, engine = 'elixir', eval = FALSE}
safeDivide(2, 0) # <= returns Left("Dividing error")
|> then(Either.map(&(&1 * 2))) # <= still Left("Dividing error")

safeDivide(2, 1) # <= returns Right(2)
|> then(Either.map(&(&1 * 2))) # <= returns Right(4)
```

The above code will use hypothetical `Left("Dividing error")` or `Right(result)`. The `Either.map/2` is a special `map` function
that runs the passed function if it's `Right` or completely ignores it when it's `Left` - it could be visualized as:

```{r, engine = 'elixir', eval = FALSE}
def map(%Either.Left{} = left, _fun), do: left
def map(%Either.Right{result: v}, fun), do: %Either.Right{result: fun.(v)}
```

This is nice and great, but what if the function inside the `Either.map/2` returns another `Either`? Like:

```{r, engine = 'elixir', eval = FALSE}
safeDivide(2, 1) # <= returns Right(2)
|> then(Either.map(&(safeDivide(&1, 1)))) # <= now Right(Right(2))!?
```

Now we need to understand those abstractions to be able to decide whether we should `map` or
`flatMap`(that's the function that will not wrap the function result into the `Right`):

```{r, engine = 'elixir', eval = FALSE}
safeDivide(2, 1) # <= returns Right(2)
|> then(Either.flatMap(&(safeDivide(&1, 1)))) <= still Right(2)
```

And that is just the beginning of the complexities that those abstractions bring.

Furthermore, let's say that inside the first `Either.map/2` callback, we will have some variable(s) that we would like to use later on.
We are now deep inside closures world like the following:

```{r, engine = 'elixir', eval = FALSE}
safeDivide(2, 1) # <= returns Right(2)
|> then(Either.flatMap(fn res ->
    # x = some data generated here
    safeDivide(2, 1)
    |> then(Either.map(&(&1 * 2)))
    |> then(Either.map(&(&1 * x))) # <= a clause to have access to x
  end))
```

The above example is obviously simplified and silly but should give us a gist of what sort of complexity we will very soon get involved in.
And, again, we just scratched the surface - there are so many more functions that the `Either` provides.
Besides, writing code in this fashion in Elixir would cause a lot of friction in the team as it's difficult to find any advantages of using it.

### Idiomatic error handling

To achieve the same results, Elixir provides the `with` statement:

```{r, engine = 'elixir', eval = FALSE}
  with {:ok, divide_result} <- safeDiv(2,1),
       {:ok, divide_result_2} <- safeDivide(2, 1)
  do
    divide_result_2 * 2 * divide_result
  else
    err -> err
  end
```

The above code provides **the same** functionality as the one before with `Either`.
We can clearly understand it **without** any knowledge about how `Either` works, `mapping`, `flatMapping` etc. It's just standard Elixir.

Again, as in the case of module attributes, Elixir provides a pragmatic way of dealing with errors -
just return a tuple with an `:error` atom. It also provides utility functions like `with` to deal with errors in an idiomatic way.
There's **no reason** to introduce concepts like `Either` as language has built-in concepts/patterns taking care of error situations.

## Do or not to do

In the last section, we discussed wrapping the results in the `Either` structs to be able to map,
flatMap on them regardless of the function result. What if we could apply the same principle to avoid executing any code(side effects) at all?

That's the basic idea behind all the category theory related abstractions like the infamous IO Monad.

I won't go into a vast amount of details.
Still, we can think about it as every time we are calling a *special* `map` or `flatMap`, instead of executing anything,
it would just wrap whatever was passed to it inside another function and return it like:

```{r, engine = 'elixir', eval = FALSE}
  def map(acc, function) do
    fn ->
      case acc.() do
        {:ok, data} -> {:ok, function.(data)}
        {:error, error} -> {:error, error}
      end
    end
  end
```

In a nutshell, what we would end up with is a function containing a function containing a function...
At this moment, I find it very difficult to find any practical reason why somebody would want to do something like this in a dynamically typed language.

In statically typed languages, there's an argument that instead of a function of function etc.,
we could have a typed object which would indicate what actions can be performed on that future result.
This is very often praised as a compile-time guarantee of side-effectful code.

In Elixir, without strong typing and with a massive impact on how the code is written and how easy it is to understand,
there's just **no practical** reason to use those concepts beyond toy programs.
The resulting function would be an untestable blob without introspection support from the BEAM VM.

## Final thoughts

**We've now separated the "what" from the "how" in our trading strategy.** Here's what we accomplished:

- Extracted pure business logic from the `Naive.Trader` into the `Naive.Strategy` module
- Split complex callbacks into single-responsibility functions that are easier to reason about
- Created decision-making functions that return data structures instead of causing side effects
- Wired up dependency injection using compile-time module attributes
- Used `with` for idiomatic error handling that's clearer than monadic abstractions

The key insight is that functional programming in Elixir isn't about avoiding side effects entirely - 
it's about knowing where to put them. Our strategy functions are now trivial to test. 
They take a state and an event, return a decision. No mocks, no setup, just pure logic.

The `Naive.Trader` still handles all the dirty work - PubSub, Binance, logging - 
but now it's a thin shell around pure decision-making. When something breaks, 
we know exactly where to look: side effects in the trader, logic in the strategy.

It's worth noting the trade-offs of this approach:

- More modules means more navigation - but each module has a clearer purpose
- Dependency injection via module attributes feels "global" - but it's the idiomatic Elixir way
- The separation introduces indirection - but it dramatically improves testability

The pattern we've established - pure logic surrounded by a thin layer of side effects - 
appears throughout idiomatic Elixir codebases. Phoenix controllers follow it: 
pure business logic in contexts, side effects (database, HTTP) at the edges.

**What About Category Theory?**

We intentionally avoided introducing monads, functors, and other category theory abstractions. 
In statically typed languages like Haskell, these abstractions provide compile-time guarantees 
about side effects. But Elixir's dynamic typing means we'd get all the complexity with none of the guarantees.

Languages like OCaml and Clojure - both functional - take the pragmatic approach we've demonstrated: 
execute side effects where needed, maximize pure functions where possible. As Rich Hickey (Clojure's creator) 
said: "It is better to have 100 functions operate on one data structure than 10 functions on 10 data structures." 
That philosophy - simple data, pure transformations - is exactly what we've built.

**So, What's Next?**

We've separated pure logic from side effects, but we haven't talked about how many 
processes we actually need. In the next chapter, we'll examine when to use processes 
for organization versus when to keep everything in one process with pure functions. 
We'll build an OHLC indicator and discover that sometimes the simplest solution 
is just... fewer processes.

[Note] Please remember to run `mix format` to keep things nice and tidy.

The source code for this chapter can be found in the book's source code repository
(branch: [chapter_17](https://github.com/Cinderella-Man/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_17))