# 80/20 win with pure logic

In the last chapter, we went in a full circle - built a complete exchange
abstraction, admitted it was premature, tore it out, and landed on `Mimic` with
compile-time module attributes. We now have a clean testing setup that doesn't
force us into unnecessary behaviours.

But we haven't exploited the real payoff yet. Remember when we split the
`Naive.Strategy` into pure functions like `generate_decision/4` and effectful
ones that talk to Binance? That separation wasn't just for aesthetics - it
was an investment. Pure functions are trivially testable: no mocks, no setup,
no state. You call them with inputs and assert on outputs.

In this chapter, we'll cash in that investment. We'll write tests for every
clause of `generate_decision/4` and show that more than half our strategy code
can be covered without touching a single external dependency. The Pareto
principle applies here almost literally - a small effort on pure code testing
buys us the majority of our confidence.

## Objectives
- testing the pure logic

## Testing the pure logic

With the testing infrastructure behind us, let's put it to work. The `generate_decision/4` function
has multiple clauses - each one a pure branch that maps inputs to a decision atom.
We'll write a test for every one of them.

We will start by removing all "hello world" tests that were generated when we were creating each of the umbrella apps - all of them look the same:

```
  test "greets the world" do
    assert $app.hello() == :world
  end
```

After taking care of this nuisance, we can now focus on testing our strategy.

Let's open the `apps/naive/test/naive/strategy_test.exs` where we will add tests of `generate_decision/4` function.
We will test each clause starting with the first (returning `:place_buy_order`):

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/test/naive/strategy_test.exs
  @tag :unit
  test "Generating place buy order decision" do
    assert Strategy.generate_decision(
             %TradeEvent{
               price: 1.0
             },
             generate_position(%{
               budget: "10.0",
               buy_down_interval: "0.01"
             }),
             :ignored,
             :ignored
           ) == {:place_buy_order, 0.99000000, "10.00000000"}
  end
```

The above test checks that the function returns `:place_buy_order` in case of a lack of buy/sell order.
Inside the test, we are using a helper function that we need to add below:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/test/naive/strategy_test.exs
  defp generate_position(data) do
    %{
      id: 1_678_920_020_426,
      symbol: "XRPUSDT",
      profit_target: "0.005",
      rebuy_interval: "0.01",
      rebuy_notified: false,
      budget: "10.0",
      buy_order: nil,
      sell_order: nil,
      buy_down_interval: "0.01",
      tick_size: "0.00010000",
      step_size: "1.00000000"
    }
    |> Map.merge(data)
    |> then(&struct(Strategy.Position, &1))
  end
```

At this moment, we should already be able to run the above test(together with the one from the previous chapter):

```
$ MIX_ENV=test mix test.unit
...
==> naive
...
2 tests, 0 failures (1 excluded)
```

We will now take care of the remaining clauses of `generate_decision/4` function.
Each test targets a specific branch: placing a sell order when a buy fills,
fetching order status when the price crosses a threshold, finishing or exiting a position,
triggering a rebuy, and the catch-all skip. The structure is identical every time -
call the function with the right inputs, assert the right atom comes back:

```{r, engine = 'elixir', eval = FALSE}
  # /apps/naive/test/naive/strategy_test.exs
  @tag :unit
  test "Generating place sell order decision" do
    assert Strategy.generate_decision(
             %TradeEvent{},
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 status: "FILLED",
                 price: 1.00
               },
               sell_order: nil,
               profit_target: "0.01",
               tick_size: "0.0001"
             }),
             :ignored,
             :ignored
           ) == {:place_sell_order, 1.0120}
  end

  @tag :unit
  test "Generating fetch buy order decision" do
    assert Strategy.generate_decision(
             %TradeEvent{
               price: 1.01
             },
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 price: 1.02
               }
             }),
             :ignored,
             :ignored
           ) == :fetch_buy_order
  end

  @tag :unit
  test "Generating finish position decision" do
    assert Strategy.generate_decision(
             %TradeEvent{},
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 status: "FILLED"
               },
               sell_order: %Binance.OrderResponse{
                 status: "FILLED"
               }
             }),
             :ignored,
             %{status: "on"}
           ) == :finished
  end

  @tag :unit
  test "Generating exit position decision" do
    assert Strategy.generate_decision(
             %TradeEvent{},
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 status: "FILLED"
               },
               sell_order: %Binance.OrderResponse{
                 status: "FILLED"
               }
             }),
             :ignored,
             %{status: "shutdown"}
           ) == :exit
  end

  @tag :unit
  test "Generating fetch sell order decision" do
    assert Strategy.generate_decision(
             %TradeEvent{
               price: 1.02
             },
             generate_position(%{
               buy_order: %Binance.OrderResponse{},
               sell_order: %Binance.OrderResponse{
                 price: 1.01
               }
             }),
             :ignored,
             :ignored
           ) == :fetch_sell_order
  end

  @tag :unit
  test "Generating rebuy decision" do
    assert Strategy.generate_decision(
             %TradeEvent{
               price: 0.89
             },
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 price: 1.00
               },
               sell_order: %Binance.OrderResponse{
                 price: 1.1
               },
               rebuy_interval: "0.1",
               rebuy_notified: false
             }),
             [:position],
             %{status: "on", chunks: 2}
           ) == :rebuy
  end

  @tag :unit
  test "Generating skip(rebuy) decision because rebuy is already notified" do
    assert Strategy.generate_decision(
             %TradeEvent{
               price: 0.89
             },
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 price: 1.00
               },
               sell_order: %Binance.OrderResponse{
                 price: 1.1
               },
               rebuy_interval: "0.1",
               rebuy_notified: true
             }),
             [:position],
             %{status: "on", chunks: 2}
           ) == :skip
  end

  @tag :unit
  test "Generating skip decision" do
    assert Strategy.generate_decision(
             %TradeEvent{
               price: 0.9
             },
             generate_position(%{
               buy_order: %Binance.OrderResponse{
                 price: 1.00
               },
               sell_order: %Binance.OrderResponse{
                 price: 1.1
               },
               rebuy_interval: "0.1",
               rebuy_notified: false
             }),
             [:position],
             %{status: "on", chunks: 1}
           ) == :skip
  end
```

This brings us to 10 tests:

```
$ MIX_ENV=test mix test.unit
...
==> naive

...
10 tests, 0 failures (1 excluded)
```

The above tests are straightforward and uneventful, but that's good. They prove that tests of pure code are easy to write and maintain.

Furthermore, besides the `generate_decision/4` function, we also have `generate_decisions/4`,
`parse_results/1` and helper methods that are all pure functions. After a little bit of math,
we can work out that out of 424 lines, 262 lines contain pure code - that's a whopping 61%.

## Final thoughts

**This was the shortest chapter in the book - and that's the point.** Here's what we covered:

- Removed boilerplate "hello world" tests generated by the umbrella scaffold
- Wrote tests for every clause of `generate_decision/4` using only pure function calls
- Measured the pure-to-effectful ratio in `Naive.Strategy` to quantify the payoff

The pattern worth remembering is the testing pyramid this creates:

1. Pure functions like `generate_decision/4` need only inputs and assertions - they're the cheapest tests to write and maintain
2. Effectful functions that call Binance or write to the database need `Mimic` stubs - more setup, but still manageable
3. Integration tests that spin up the full system are the most expensive and should cover the least ground

The underlying principle is simple: push side effects to the edges, keep the core pure, and your test suite becomes fast,
reliable, and cheap to maintain. We've been doing exactly that - perhaps without naming it -
since we split `Naive.Strategy` into decision functions and execution functions.

**So, What's Next?**

Our strategy is tested, our mocking story is clean, and our code is structured
well. But we're still running as an umbrella application with multiple databases
and duplicated configuration. In the next chapter, we'll ditch the umbrella
entirely and migrate to a single Phoenix application - simplifying deployment
and setting us up for clustering.

[Note] Please remember to run `mix format` to keep things nice and tidy.

The source code for this chapter can be found in the book's source code repository
(branch:
[chapter_21](https://github.com/Cinderella-Man/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_21)).
