Chapter 23

# Back to the Monolith

## Objectives
- reverting the course to the monolith
- creating a new Phoenix app
- reintegrating the 'core' app
- reintegrating the `binance_mock` app
- reintegrating the `streamer` app
- reintegrating the `naive` app

## Reverting the course to the monolith

We started our journey with the idea that we would use an umbrella to "future-proof" our project in case we would migrate to microservices architecture.

As we delved deeper into our project(which also included me gaining more knowledge and experience over three years of writing this book), we began to see that the umbrella approach was, at best, an example of overengineering.

Looking back, the umbrella approach introduced significant complexity, a factor that became increasingly difficult to justify as our understanding of the project evolved.

For example, each umbrella app has its configuration, which resulted in using multiple databases and duplicated settings for things like the Binance client, making our deployment unnecessarily expensive and complex.

In this chapter, we will simplify the whole project by ditching the umbrella and migrating to Phoenix, which will be instrumental in our next steps, including deployment and clustering.

## Creating a new Phoenix app

Instead of improving the existing structure, we will start from scratch. We will create a new Phoenix project side-by-side next to the current umbrella app, and then we will reintegrate each nested app one after another.

First, let's make sure that we have the latest version of the Phoenix application generator installed:

```
$ mix archive.install hex phx_new
```

Note: At the time of writing this chapter, the current version of the Phoenix framework is 1.7.12.

With this taken care of, we can progress with the creation of a new Phoenix application and a database:

```
$ mix phx.new hedgehog
...
$ cd hedgehog
$ mix ecto.create
```

With the Phoenix app skeleton in place, we can move on to reintegrating the umbrella apps' contents into it.

## Reintegrating the 'core' app

We will start with the `core` app, as it is used/referenced by other apps in the umbrella.

It contains just a single module called `Core.Struct.TradeEvent` (inside the `apps/core/lib/core/struct/trade_event.ex` file) which we will rename to `Hedgehog.Exchange.TradeEvent` inside the newly created Phoenix app (we need to create a new `exchange` directory inside the `/lib/hedgehog` and paste the `trade_event.ex` file inside).

Besides the `TradeEvent` module, the `core` application supervises the `PubSub` supervision tree. After a quick check inside the `/lib/hedgehog/application.ex` file, we can confirm that Phoenix already uses PubSub:

```{r, engine = 'elixir', eval = FALSE}
  # /lib/hedgehog/application.ex
    children = [
      ...
      {Phoenix.PubSub, name: Hedgehog.PubSub},
      ...
    ]
```

Furthermore, the `phoenix_pubsub` dependency is already included in the `mix.lock` file, and its default backend is now `Phoenix.PubSub.PG2`(the adapter we have been explicitly specifying before).

The above change finishes merging the `core` app into our new Phoenix app. We should be able to run our app:

```
$ iex -S mix phx.server
...
[info] Running HedgehogWeb.Endpoint with Bandit 1.4.2 at 127.0.0.1:4000 (http)
...
iex(1)> alias Hedgehog.Exchange.TradeEvent
...
iex(2)> %TradeEvent{}
%Hedgehog.Exchange.TradeEvent{
...
}
```

## Reintegrating the `binance_mock` app

The `binance_mock` app is used by both the `naive` and `streamer` apps, so we will need to focus on it next.

First, we will find a new home for the cached exchange info file which was previously based in `/apps/binance_mock/test/assets/exchange_info.json`. We will create a new directory called `/priv/cache` where we will paste the `exchange_info.json` file.

Next, we will take care of configuration that was deciding should we use the cached exchange info file mentioned above - inside the new application it will look as follows:

```{r, engine = 'elixir', eval = FALSE}
  # /config/config.exs
config :hedgehog,
  ecto_repos: [Hedgehog.Repo],
  generators: [timestamp_type: :utc_datetime],
  exchanges: [ # <= added
    binance_mock: [ 
      use_cached_exchange_info: true
    ]
  ]
```

Moving on to the main file of interest - `apps/binance_mock/lib/binance_mock.ex`, which we will move to the `/lib/hedgehog/exchange` directory.

Inside the module, we need to apply the following changes:
change the module name to `Hedgehog.Exchange.BinanceMock`
update the alias to the `Core.Struct.TradeEvent` struct to be `Hedgehog.Exchange.TradeEvent`
update references to the `Core.PubSub` with `Hedgehog.PubSub`
extract to boolean config flag to a module's attribute and use it inside the `get_exchange_info/0` function:
```{r, engine = 'elixir', eval = FALSE}
  # /lib/hedgehog/exchange/binance_mock.ex
  @use_cached_exchange_info Application.compile_env!(:hedgehog, [
                              :exchanges,
                              :binance_mock,
                              :use_cached_exchange_info
                            ])
  ...
  def get_exchange_info() do
    case @use_cached_exchange_info do
    ...
```
update the `get_cached_exchange_info/0` function to point to the new location of the `exchange_info.json` file:
```{r, engine = 'elixir', eval = FALSE}
  # /lib/hedgehog/exchange/binance_mock.ex    
  defp get_cached_exchange_info do
    {:ok, data} =
      File.cwd!()
      |> Path.split()
      |> Kernel.++([
        "priv",
        "cache",
        "exchange_info.json"
      ])
      |> Path.join()
      |> File.read()
    ...
```

That finishes our changes to the `Hedgehog.Exchange.BinanceMock` module which we need to add supervision tree of our application:

```{r, engine = 'elixir', eval = FALSE}
  # /lib/hedgehog/application.ex
  def start(_type, _args) do
    children = [
      ...
      Hedgehog.Exchange.BinanceMock
    ...
```

The `Hedgehog.Exchange.BinanceMock` module depends on a few packages that we need to add to the `mix.exs` dependencies:

```{r, engine = 'elixir', eval = FALSE}
  # /mix.exs
  defp deps do
    [
      ...
      {:binance, "~> 1.0"},
      {:decimal, "~> 2.0"},
      ...
    ]
```

Please remember to run the `mix deps.get` before trying out our changes:

```
$ iex -S mix phx.server
...
iex(1)> alias Hedgehog.Exchange.BinanceMock
iex(2)> Process.whereis(BinanceMock)
#PID<...>
iex(3)> |> Process.alive?()
true
iex(4)> BinanceMock.get_exchange_info()
%{
  ...
}
```

The above confirms that we now have a `BinanceMock` process running in the background and we are able to fetch exchange info.

## Reintegrating the `streamer` app

In the case of the `streamer` app, there are multiple files to move, so we will first create a new `/lib/hedgehog/streamer`(inside the new Phoenix app) directory and then copy both the `/apps/streamer/lib/streamer` directory and the `/apps/streamer/lib/streamer.ex` file into that new directory.

As we will be updating/renaming all of those files, we will use this opportunity to place all of them inside the `Binance` namespace.

We now need to modify each file starting with renaming `/lib/hedgehog/streamer/streamer.ex` to `/lib/hedgehog/streamer/binance.ex` and updating the module name and alias:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance.ex
defmodule Hedgehog.Streamer.Binance do
  ...
  alias Hedgehog.Streamer.Binance.DynamicStreamerSupervisor
```

The next step will be to update the `/lib/hedgehog/streamer/streamer` directory to `/lib/hedgehog/streamer/binance`. We will now move on to the files inside this directory.

### Supervisor

Starting with the `supervisor.ex` file, we need to update the module's name and alias:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance/supervisor.ex
defmodule Hedgehog.Streamer.Binance.Supervisor do
   ...
   alias Hedgehog.Streamer.Binance.DynamicStreamerSupervisor
```

Moving forward, we can remove the `repo.ex` file as we will use the `Hedgehog.Repo` module to work with the database.

### Worker

Next, we will rename the `/lib/hedgehog/streamer/binance/binance.ex` to  `/lib/hedgehog/streamer/binance/worker.ex` - mainly to avoid "binance/binance" name after we added the namespace. Following the filename change, we need to update the module's name and a couple of references to the `Core` module:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance/worker.ex
defmodule Hedgehog.Streamer.Binance.Worker do
...
  defp process_event(%{"e" => "trade"} = event) do
    trade_event = %Hedgehog.Exchange.TradeEvent{ # <= updated
...
    Phoenix.PubSub.broadcast(
      Hedgehog.PubSub, # <= updated
...
```

### DynamicStreamerSupervisor

Next, we will update the `dynamic_streamer_supervisor.ex`, where we will update the module's name and all the aliases:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance/dynamic_streamer_supervisor.ex
defmodule Hedgehog.Streamer.Binance.DynamicStreamerSupervisor do
...
  alias Hedgehog.Repo
  alias Hedgehog.Streamer.Binance.Worker
  alias Hedgehog.Streamer.Settings
...
  defp start_child(args) do
    DynamicSupervisor.start_child(
      __MODULE__,
      {Worker, args} # <= updated
    )
  end
...
```

### schema/settings.ex and schema/streaming_status_enum.ex

We will move the `/lib/hedgehog/streamer/binance/schema/settings.ex` file to `/lib/hedgehog/streamer/settings.ex` and the `/lib/hedgehog/streamer/binance/schema/streaming_status_enum.ex` file to 
`/lib/hedgehog/streamer/settings_status_enum.ex`. We can now remove the empty `/lib/hedgehog/streamer/binance/schema` directory. Both of those modules need their modules' names and references updated:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/settings.ex
defmodule Hedgehog.Streamer.Settings do
  use Ecto.Schema

  alias Hedgehog.Streamer.SettingsStatusEnum

  @primary_key {:id, :binary_id, autogenerate: true}

  schema "streaming_settings" do
    field(:symbol, :string)
    field(:status, SettingsStatusEnum)

    timestamps()
  end
end
```

and:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/settings_status_enum.ex
import EctoEnum

defenum(Hedgehog.Streamer.SettingsStatusEnum, :status, [:on, :off])
```

### application

The final file that we need to look into is `/lib/hedgehog/streamer/binance/application.ex`, where we will look into children who were supervised by the `streamer` app:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance/application.ex
    children = [
      {Streamer.Repo, []},
      {Streamer.Supervisor, []}
    ]
```

We don't need to worry about the `Streamer.Repo` as we will use `Hedgehog.Repo`. On the other hand, the `Streamer.Supervisor` which we renamed to `Hedgehog.Streamer.Binance.Supervisor` needs to be added to the main `Hedgehog.Application` module:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/application.ex
    children = [
      ...
      Hedgehog.Streamer.Binance.Supervisor
    ]
```

We can now remove `/lib/hedgehog/streamer/binance/application.ex` as it is no longer required.

### DB migrations and seeding

 Streaming depends on the database tables and settings(seed data). We will start by copying `apps/streamer/priv/repo/migrations/20210203184805_create_settings.exs` migration to `/priv/repo/migrations`(and renaming the file to `20210203184805_create_streamer_settings.ex`) directory and update the most of the code inside:

```{r, engine = 'elixir', eval = FALSE}
# /priv/repo/migrations/20210203184805_create_streamer_settings.exs
defmodule Hedgehog.Repo.Migrations.CreateStreamerSettings do
  use Ecto.Migration

  alias Hedgehog.Streamer.SettingsStatusEnum

  def change do
    SettingsStatusEnum.create_type()

    create table(:streamer_settings, primary_key: false) do
      add(:id, :uuid, primary_key: true)
      add(:symbol, :text, null: false)
      add(:status, SettingsStatusEnum.type(), default: "off", null: false)

      timestamps()
    end

    create(unique_index(:streamer_settings, [:symbol]))
  end
end
```

Next, we will copy the seeding script's code from `/apps/streamer/priv/seed_settings.exs` to `/priv/repo/seeds.exs` and make the following updates:

```{r, engine = 'elixir', eval = FALSE}
# /priv/repo/seeds.exs
alias Hedgehog.Repo
alias Hedgehog.Streamer.Settings

binance_client = Application.compile_env(:hedgehog, :binance_client)
```

### Config

Inside the script above, we were reading the application's configuration expecting the `binance_client` to be there - let's append the required settings to the configuration:

```{r, engine = 'elixir', eval = FALSE}
# /config/config.exs
config :hedgehog,
  binance_client: Hedgehog.Exchange.BinanceMock, # <= added
  ecto_repos: [Hedgehog.Repo],
  ...
```

### Deps

We need to add a couple of new dependencies that the streaming code is using:

 ```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  defp deps do
    [
      ...
      {:ecto_enum, "~> 1.4"},
      {:websockex, "~> 0.4.2"}
```

With the above changes, the reintegration of the `Streamer` app is finished. We need to remember about getting new deps, running migrations, seeding database and we can test that everything works up to this point:

```
$ mix deps.get
...
$ mix ecto.migrate
...
$ mix run priv/repo/seeds.exs
...
$ iex -S mix phx.server
...
iex(1)> Hedgehog.Streamer.Binance.start_streaming("XRPUSDT")
...
[info] Binance streamer is connecting to websocket stream for XRPUSDT trade events
{:ok, #PID<0.801.0>}
[debug] Trade event received XRPUSDT@0.55080000
BREAK: (a)bort (A)bort with dump (c)ontinue (p)roc info (i)nfo
       (l)oaded (v)ersion (k)ill (D)b-tables (d)istribution
^C
$ iex -S mix phx.server
...
[info] Binance streamer is connecting to websocket stream for XRPUSDT trade events
[debug] Trade event received XRPUSDT@0.55240000
iex(1)> Hedgehog.Streamer.Binance.stop_streaming("XRPUSDT")
[info] Stopping streaming XRPUSDT trade events
...
```

The above confirms that we start and stop streaming, and upon startup, streaming starts on its own when enabled in the database.

## Reintegrating the `naive` app

++ code	
++ config
++ migrations
++ seed
++ tests
