# From Postgres to SQLite

In the last chapter, we consolidated six umbrella apps into a single Phoenix
application - one repo, one database, one config surface. The structure is clean.
But if you look at how it runs, there's still an external process standing between
our application and its data: Postgres, managed via Docker.

For development, that means running `docker compose up -d` before you can do
anything. For deployment, it means each trading node needs a network path to
a shared Postgres instance - and each write is a round-trip over that link.
If the database goes down or the network hiccups, all nodes stop trading.

For a system we intend to distribute across independent nodes, this is exactly
the wrong architecture. We want each node to be self-contained: its own
processes, its own state, its own persistence.

SQLite gives us that. It's an embedded database that lives as a single file
alongside the application - no server process, no network dependency, no Docker container.
If that sounds like a toy database, it isn't.

Benchmarks routinely show SQLite handling over 100k SELECTs per second on modest hardware
with proper tuning - WAL mode, memory-mapped I/O, and normal synchronous mode
(see phiresky's "SQLite performance tuning" at phiresky.github.io for the full breakdown).

For a single-node writer like our trading bot, it's more than enough.

The `ecto_sqlite3` adapter makes the switch almost transparent to our application
code; Ecto's Repo API stays the same, migrations need only minor adjustments.

In this chapter, we'll swap Postgres for SQLite, remove the Docker dependency
entirely, and replace the `ecto_enum` library(which relies on Postgres-specific
`CREATE TYPE` statements) with Ecto's built-in `Ecto.Enum`. We'll also set up
Litestream for continuous database backups to S3-compatible storage - because an
embedded database without off-site replication is one disk failure away from total
data loss - and wire up automatic restore so a fresh node recovers its database
from backup before the first trade. By the end, our trading bot will be a fully
self-contained, self-healing application that carries its database with it - which
is exactly what we'll need when we introduce clustering in the next chapter.

## Objectives
- removing the Docker/Postgres dependency
- swapping the database adapter to SQLite
- replacing `ecto_enum` with `Ecto.Enum`
- updating status values from strings to atoms
- updating migrations for SQLite compatibility
- setting up continuous backups with Litestream and S3-compatible storage
- automating database recovery on startup

## Removing the Docker dependency

The `docker-compose.yml` file at the root of our project exists solely to run
Postgres. Since we're moving to SQLite, we no longer need it:

```{r, engine = 'bash', eval = FALSE}
rm docker-compose.yml
```

We should also add SQLite's database files to `.gitignore`. SQLite uses up to
three files per database (the main `.db` file, a write-ahead log `-wal`, and a
shared-memory file `-shm`):

```{r, engine = 'text', eval = FALSE}
# /.gitignore
# Database files
*.db
*.db-*
backups/
```

## Swapping the database adapter

### Dependencies

Inside `mix.exs`, we need to swap `postgrex` for `ecto_sqlite3` and remove
`ecto_enum`(we'll replace it with Ecto's built-in `Ecto.Enum` in the next section):

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  defp deps do
    [
      ...
      {:ecto_sqlite3, "~> 0.22"}, # <= added (replaces postgrex)
      ...
    ]
  end
```

Remove the following lines from `deps`:

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  defp deps do
    [
      ...
      {:postgrex, ">= 0.0.0"},  # <= remove
      ...
      {:ecto_enum, "~> 1.4"},   # <= remove
      ...
    ]
  end
```

### Repo adapter

Next, we need to update `Hedgehog.Repo` to use the SQLite3 adapter:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/repo.ex
defmodule Hedgehog.Repo do
  use Ecto.Repo,
    otp_app: :hedgehog,
    adapter: Ecto.Adapters.SQLite3 # <= updated
end
```

### Configuration

The Postgres configuration used hostname, username, password, and a database name.
SQLite only needs a file path. We need to update all environment config files.

Starting with the development configuration:

```{r, engine = 'elixir', eval = FALSE}
# /config/dev.exs
config :hedgehog, Hedgehog.Repo,
  database: Path.expand("../hedgehog_dev.db", Path.dirname(__ENV__.file)),
  pool_size: 5,
  show_sensitive_data_on_connection_error: true
```

SQLite supports multiple concurrent readers in WAL mode, so a pool size of 5 is a
reasonable default - enough to allow several read queries to overlap.
Writes are serialized regardless of pool size; SQLite permits only one writer at a time.

The `ecto_sqlite3` adapter relies on SQLite's built-in busy-wait mechanism to handle write
contention - if a connection attempts to write while another write is in progress, it will
retry for up to `busy_timeout` milliseconds (default: 2000ms) before raising an error.

A larger pool won't speed up writes, but it can help if you have many concurrent reads.

**Note:** `Path.expand("../hedgehog_dev.db", Path.dirname(__ENV__.file))` resolves
to a file in the project root directory. This is the standard convention for
Phoenix projects using SQLite.

The test configuration follows the same pattern but uses a separate database file:

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
config :hedgehog, Hedgehog.Repo,
  database: Path.expand("../hedgehog_test.db", Path.dirname(__ENV__.file)),
  pool_size: 5,
  pool: Ecto.Adapters.SQL.Sandbox
```

**Note:** The `ecto_sqlite3` adapter does not support async tests with
`Ecto.Adapters.SQL.Sandbox`. If you have test modules marked with
`async: true`, they'll need to run synchronously when using SQLite.
This is a known limitation caused by SQLite's single-writer constraint -
the Sandbox wraps each test in a transaction, and concurrent write
transactions deadlock.

And the integration configuration:

```{r, engine = 'elixir', eval = FALSE}
# /config/integration.exs
config :hedgehog, Hedgehog.Repo,
  database: Path.expand("../hedgehog_integration.db", Path.dirname(__ENV__.file)),
  pool_size: 5,
  pool: Ecto.Adapters.SQL.Sandbox
```

Finally, the runtime configuration for production. We replace the `DATABASE_URL`
environment variable with `DATABASE_PATH`:

```{r, engine = 'elixir', eval = FALSE}
# /config/runtime.exs
if config_env() == :prod do
  database_path =
    System.get_env("DATABASE_PATH") ||
      raise """
      environment variable DATABASE_PATH is missing.
      For example: /data/hedgehog/hedgehog.db
      """
  # remove: maybe_ipv6 = ....

  config :hedgehog, Hedgehog.Repo,
    database: database_path,
    pool_size: String.to_integer(System.get_env("POOL_SIZE") || "5")
  ...
```

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note regarding releases:** In dev.exs, we used `Path.expand` relative to the current file.
In `runtime.exs` (Production), we rely on the `DATABASE_PATH` environment variable.
This is critical because when you build an OTP Release,
the configuration files are evaluated, but the source code directories may not exist on the target server.
Always use absolute paths or environment variables for database locations in production.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on concurrency:** As described above, SQLite serializes writes and any blocked
writer waits up to `busy_timeout` milliseconds before raising an error.

For a trading bot where multiple GenServers may write to the database, you might want to
increase this value beyond the default 2000ms. Our current setup is fine with the default,
but if you add more concurrent writers later, bumping it prevents spurious timeouts:

```{r, engine = 'elixir', eval = FALSE}
# /config/*.exs
    config :hedgehog, Hedgehog.Repo,
      ...
      busy_timeout: 5000
```

See the `ecto_sqlite3` adapter documentation for the full list of tuning options.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

At this point, our adapter is configured but our migrations and schemas still
reference `ecto_enum` types that only work with Postgres. Let's fix that next.

## Replacing `ecto_enum` with `Ecto.Enum`

This is the most involved part of the migration - not because it's difficult,
but because the `ecto_enum` library is woven into three separate areas of our
codebase: schema files, migration files, and dedicated enum modules.

### Why the change is necessary

The `ecto_enum` library uses Postgres-specific `CREATE TYPE` statements to define
custom enum types at the database level. SQLite has no equivalent feature - it
stores everything as one of five types: NULL, INTEGER, REAL, TEXT, or BLOB.

Ecto's built-in `Ecto.Enum` takes a different approach:
it stores enum values as plain strings in the database and handles the atom
conversion in the application layer. This works identically across all database
backends.

### Removing the enum modules

We have three `*_status_enum.ex` files that use `defenum` from the `ecto_enum` library.
Since `Ecto.Enum` is declared inline within schema definitions, these dedicated modules are
no longer needed. We can delete all three:

```{r, engine = 'bash', eval = FALSE}
rm lib/hedgehog/streamer/settings_status_enum.ex
rm lib/hedgehog/strategy/naive/settings_status_enum.ex
rm lib/hedgehog/data/collector/settings_status_enum.ex
```

### Updating the schema files

Each settings schema currently references its companion enum module. We need to
replace those references with `Ecto.Enum` field declarations.

Starting with the streamer settings:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/settings.ex
defmodule Hedgehog.Streamer.Settings do
  use Ecto.Schema

  # remove: alias Hedgehog.Streamer.SettingsStatusEnum

  ...

  schema "streamer_settings" do
    ...
    field(:status, Ecto.Enum, values: [:on, :off]) # <= updated
    ...
  end
end
```

Next, the naive strategy settings - note that this enum has three values
including `:shutdown`:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/strategy/naive/settings.ex
defmodule Hedgehog.Strategy.Naive.Settings do
  use Ecto.Schema

  # remove: alias Hedgehog.Strategy.Naive.SettingsStatusEnum

  ...

  schema "naive_strategy_settings" do
    ...
    field(:status, Ecto.Enum, values: [:on, :off, :shutdown]) # <= updated
    ...
  end
end
```

And finally, the collector settings:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/data/collector/settings.ex
defmodule Hedgehog.Data.Collector.Settings do
  use Ecto.Schema

  # remove: alias Hedgehog.Data.Collector.SettingsStatusEnum

  ...

  schema "collector_settings" do
    ...
    field(:status, Ecto.Enum, values: [:on, :off]) # <= updated
    ...
  end
end
```

### Important: strings must become atoms

There is one catch. The old `ecto_enum` library was lenient - it happily accepted
both strings (`"on"`) and atoms (`:on`). `Ecto.Enum` is stricter: it only
accepts atoms. Any place in our codebase that passes a string like `"on"` or
`"off"` to an Ecto query or an `insert_all` call will now raise an `Ecto.ChangeError` at runtime.

We need to find and update every occurrence. There are three areas to fix:
autostart queries, status update functions, and the seed script.

### Autostart queries

Three supervisor modules query for settings `where: s.status == "on"`.
We need to change the string `"on"` to the atom `:on` in each:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance/dynamic_streamer_supervisor.ex
  def autostart_workers do
    Repo.all(
      from(s in Settings,
        where: s.status == :on, # <= updated from "on" to :on
        select: s.symbol
      )
    )
    ...
```

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/strategy/naive/dynamic_trader_supervisor.ex
  def autostart_workers do
    Repo.all(
      from(s in Settings,
        where: s.status == :on, # <= updated from "on" to :on
        select: s.symbol
      )
    )
    ...
```

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/data/collector/dynamic_worker_supervisor.ex
  def autostart_workers do
    Repo.all(
      from(s in Settings,
        where: s.status == :on, # <= updated from "on" to :on
        select: s.topic
      )
    )
    ...
```

### Status update functions

Several modules have `update_status/2` functions that pass string status values
when building changesets or structs. We need to update these to atoms as well:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/streamer/binance/dynamic_streamer_supervisor.ex
  defp update_status(symbol, status)
       when is_binary(symbol) and status in [:on, :off] do # <= updated
    ...
```

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/data/collector/dynamic_worker_supervisor.ex
  defp update_status(topic, status)
       when is_binary(topic) and status in [:on, :off] do # <= updated
    ...
```

In the naive strategy, the Formula module also references status values:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/strategy/naive/formula.ex
  def update_status(symbol, status)
      when is_binary(symbol) and status in [:on, :off, :shutdown] do # <= updated
    ...
```

We also need to update all the *callers* of these functions. Search for any
call sites that pass string arguments like `update_status(symbol, "on")` and
replace them with atoms:

- `update_status(..., "on")` becomes `update_status(..., :on)`
- `update_status(..., "off")` becomes `update_status(..., :off)`
- `update_status(..., "shutdown")` becomes `update_status(..., :shutdown)`

These calls appear in `start_worker`/`stop_worker` and `start_child`/`stop_child`
functions across the three dynamic supervisor modules and the formula module.

### Updating the seed script

The seeding script in `/priv/repo/seeds.exs` also uses string values for status
fields inside the maps passed to `Repo.insert_all`. Since `Ecto.Enum` requires
atoms, we need to update both the streamer and naive settings sections:

```{r, engine = 'elixir', eval = FALSE}
# /priv/repo/seeds.exs
# In the streamer base_settings map:
base_settings = %{
  ...
  status: :off, # <= updated from "off" to :off
  ...
}
...
# In the naive strategy base_settings map:
base_settings = %{
  ...
  status: :off, # <= updated from "off" to :off
  ...
}
```

## Updating migrations for SQLite compatibility

Our existing migrations use two Postgres-specific features that SQLite doesn't
support: custom enum types (via `create_type()`) and `ALTER TYPE` statements.
We need to rewrite them to use plain `:string` columns instead.

### Streamer settings migration

The original migration calls `SettingsStatusEnum.create_type()` and references
`SettingsStatusEnum.type()`. We'll replace both with a simple `:string` column:

```{r, engine = 'elixir', eval = FALSE}
# /priv/repo/migrations/20210203184805_create_streamer_settings.exs
defmodule Hedgehog.Repo.Migrations.CreateStreamerSettings do
  use Ecto.Migration

  # remove: alias Hedgehog.Streamer.SettingsStatusEnum

  def change do
    # remove: SettingsStatusEnum.create_type()

    create table(:streamer_settings, primary_key: false) do
      ...
      add(:status, :string, default: "off", null: false) # <= updated
      ...
    end
    ...
  end
end
```

### Naive strategy settings migration

Same pattern - remove the enum type creation and use `:string`:

```{r, engine = 'elixir', eval = FALSE}
# /priv/repo/migrations/20210202223209_create_naive_strategy_settings.exs
defmodule Hedgehog.Repo.Migrations.CreateNaiveStrategySettings do
  use Ecto.Migration

  # remove: alias Hedgehog.Strategy.Naive.SettingsStatusEnum

  def change do
    # remove: SettingsStatusEnum.create_type()

    create table(:naive_strategy_settings, primary_key: false) do
      ...
      add(:status, :string, default: "off", null: false) # <= updated
      ...
    end
    ...
  end
end
```

### Naive strategy status update migration

The second naive migration used `ALTER TYPE` to add the `:shutdown` value to the
Postgres enum. Since we're now storing statuses as plain strings, there's no type
to alter - the column already accepts any string value, and `Ecto.Enum` handles
validation in the application layer.

Since we're working from a clean database(SQLite is brand new), we can safely
remove this migration entirely:

```{r, engine = 'bash', eval = FALSE}
rm priv/repo/migrations/20210205232303_update_naive_strategy_settings_status.exs
```

### Collector settings migration

```{r, engine = 'elixir', eval = FALSE}
# /priv/repo/migrations/20210227230123_create_collector_settings.exs
defmodule Hedgehog.Repo.Migrations.CreateCollectorSettings do
  use Ecto.Migration

  # remove: alias Hedgehog.Data.Collector.SettingsStatusEnum

  def change do
    # remove: SettingsStatusEnum.create_type()

    create table(:collector_settings, primary_key: false) do
      ...
      add(:status, :string, default: "off", null: false) # <= updated
      ...
    end
    ...
  end
end
```

Let's verify the script works by running the full setup after fetching
our new dependencies:

```{r, engine = 'bash', eval = FALSE}
mix deps.get
...
mix ecto.create
...
mix ecto.migrate
...
mix run priv/repo/seeds.exs
...
[info] Inserting default streamer settings for symbols
[info] Inserted batch of 1000 streamer settings
...
[info] Inserting default naive strategy settings for symbols
[info] Inserted batch of 1000 naive settings
...
```

Note: Unlike Postgres, SQLite creates the database file automatically the first time anything opens it -
including `mix ecto.migrate`. This means `mix ecto.create` is effectively optional with SQLite.
If it complains about an existing database, you can safely ignore the warning or run `mix ecto.drop` first to start fresh.

## Testing

Let's run the full test suite to make sure the migration didn't break anything:

```{r, engine = 'bash', eval = FALSE}
mix test.unit
..........
Finished in 0.1 seconds (0.1s async, 0.00s sync)
10 tests, 0 failures (6 excluded)
```

```{r, engine = 'bash', eval = FALSE}
MIX_ENV=integration mix test.integration
..
Finished in 7.0 seconds (0.05s async, 7.0s sync)
1 test, 0 failures (15 excluded)
```

**Note:** ExUnit always reports `async` and `sync` timing even when no tests are marked `async: true`.
The label reflects ExUnit's internal scheduling, not SQLite concurrency.

And a manual smoke test to verify the full trading cycle still works:

```{r, engine = 'bash', eval = FALSE}
mix ecto.drop
...
mix setup
...
iex -S mix phx.server
...
iex(1)> alias Hedgehog.Streamer.Binance
...
iex(2)> alias Hedgehog.Strategy.Naive
...
iex(3)> alias Hedgehog.Data.Collector
...
iex(4)> Collector.start_storing("TRADE_EVENTS", "XRPUSDT")
...
iex(5)> Collector.start_storing("ORDERS", "XRPUSDT")
...
iex(6)> Binance.start_streaming("XRPUSDT")
...
iex(7)> Naive.start_trading("XRPUSDT")
...
[info] Position (XRPUSDT/1770670753113): Placing a BUY order @ 1.455, quantity: 137.40000000
...
[info] Position (XRPUSDT/1770670753113): The BUY order is now filled
...
[info] Position (XRPUSDT/1770670753113): Placing a SELL order @ 1.4561, quantity: 137.40000000
...
[info] Position (XRPUSDT/1770670753113): The SELL order is now filled
```

Everything works exactly as before - streaming, trading, collecting - but now
there's no Docker container running in the background. The database is just a
file sitting in our project root:

```{r, engine = 'bash', eval = FALSE}
ls -la hedgehog_dev.db*
-rw-r--r-- 1 user user 442368 Feb 11 14:23 hedgehog_dev.db
-rw-r--r-- 1 user user  32768 Feb 11 14:23 hedgehog_dev.db-shm
-rw-r--r-- 1 user user 524288 Feb 11 14:23 hedgehog_dev.db-wal
```

## Backing up the database

With Postgres, backups are a well-trodden path - `pg_dump`, WAL archiving,
managed hosting that handles it for you.
With SQLite, the database is a file sitting in our project root.
That simplicity is the whole point - but it also means backups are
entirely our responsibility.

You might think `cp hedgehog_dev.db hedgehog_backup.db` would do the trick, but it won't.
If the application is running, a raw file copy is unreliable. In WAL mode the database state
is split across the main `.db` file and the `-wal` file. Copying just the `.db` file gives you
a valid but stale snapshot(missing any changes still in the WAL).
Copying both files at different moments can produce an inconsistent backup.
Even if you could copy them atomically, a local backup on the same disk isn't really a backup - if the disk fails, you lose both.

SQLite does provide safe one-shot backup mechanisms(`VACUUM INTO` or the `.backup`
command), but those produce point-in-time snapshots. For a trading bot where every
executed trade matters, losing hours of data to a periodic snapshot is not acceptable.
What we need is continuous replication to remote storage.

Litestream does exactly this. It monitors the SQLite database and streams every
committed transaction to S3-compatible storage using the LTX file format - a
transaction-aware format that supports compaction, making restores significantly
faster than replaying individual WAL segments. By default, Litestream syncs new
changes every 1 second, giving you an RPO(Recovery Point Objective) of roughly
one second. This is comparable to the replication lag you'd see with Postgres
asynchronous streaming replication, which is the mode most production Postgres
setups actually run.

The trade-off is recovery time. With a Postgres hot standby, failover takes seconds -
the replica is already running and can be promoted immediately. With SQLite + Litestream,
recovering from a dead node means provisioning a new server and starting the application -
we'll wire up automatic restore later in this chapter so the app pulls the database from
S3 before the Repo boots. For a database under a few hundred megabytes like ours, that's
a couple of minutes at most.
It's worth knowing that you're trading faster failover for simpler operations.

The `litestream` Hex package wraps the Litestream binary in a supervised GenServer. It
downloads the binary automatically on first start, manages it as an Erlang port, and
integrates cleanly with our supervision tree. Let's set it up.

### Adding the dependency

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  defp deps do
    [
      ...
      {:litestream, "~> 1.0"}, # <= added
      ...
    ]
  end
```

Fetch the new dependency:

```{r, engine = 'bash', eval = FALSE}
mix deps.get
...
```

### Configuring Litestream

Litestream needs to know which Repo to back up, where to replicate it, and how to
authenticate. We'll configure these per environment.

For development and test, we disable Litestream entirely - there's no reason to
replicate a throwaway dev database to remote storage:

```{r, engine = 'elixir', eval = FALSE}
# /config/dev.exs
config :hedgehog, Hedgehog.Litestream, enabled: false
```

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
config :hedgehog, Hedgehog.Litestream, enabled: false
```

```{r, engine = 'elixir', eval = FALSE}
# /config/integration.exs
config :hedgehog, Hedgehog.Litestream, enabled: false
```

For production, the credentials and storage details come from environment variables.
Add the following inside the existing `if config_env() == :prod do` block in `runtime.exs`:

```{r, engine = 'elixir', eval = FALSE}
# /config/runtime.exs (inside the :prod block)
  replica_url =
    System.get_env("LITESTREAM_REPLICA_URL") ||
      raise """
      environment variable LITESTREAM_REPLICA_URL is missing.
      For example: https://<account-id>.r2.cloudflarestorage.com
      """

  node_name =
    System.get_env("HEDGEHOG_NODE_NAME") ||
      System.get_env("RELEASE_NODE") ||
      raise """
      environment variable HEDGEHOG_NODE_NAME or RELEASE_NODE is missing.
      This identifies the node for backup isolation.
      For example: hedgehog_us_east_1
      """

  config :hedgehog, Hedgehog.Litestream,
    enabled: true,
    repo: Hedgehog.Repo,
    endpoint: replica_url,
    bucket: "hedgehog",
    replica_path: "backups/#{node_name}",
    region: "auto",
    override_architecture: :x86_64,
    access_key_id: System.fetch_env!("LITESTREAM_ACCESS_KEY_ID"),
    secret_access_key: System.fetch_env!("LITESTREAM_SECRET_ACCESS_KEY")
```

There are a few more configuration options here than you might expect. The Litestream
binary's CLI only accepts a subset of replica settings as command-line arguments -
options like `region` and replica `type` can only be specified through a YAML config file.
The `litestream` package handles this transparently by generating an ad-hoc YAML
configuration file at runtime from these Elixir options, then passing it to the binary.
This is what enables S3-compatible services like Cloudflare R2 that require specific
region and type settings to work correctly.

Notice the `replica_path` uses the node name as a prefix. This is critical for
multi-node deployments: each trading node writes to its own path inside the shared
bucket (e.g. `backups/hedgehog_us_east_1/`, `backups/hedgehog_eu_west_1/`). Without
this, two nodes replicating to the same path would overwrite each other's backup
state - and restoring would give you a Frankenstein database mixing data from
different nodes.

We use `HEDGEHOG_NODE_NAME` when set explicitly, falling back to
`RELEASE_NODE` which Elixir's release system provides automatically. Prefer setting
`HEDGEHOG_NODE_NAME` to a stable identifier - `RELEASE_NODE` often includes the
hostname(e.g. `hedgehog@10.0.1.5`), which can change across container restarts and
would make restore paths unpredictable.

Let's go through the options:

- **`repo`** - the Ecto Repo whose database path Litestream should watch
- **`endpoint`** - the S3-compatible endpoint URL (just the bare endpoint, not an S3 URI)
- **`bucket`** - the name of the storage bucket
- **`replica_path`** - a key prefix inside the bucket, scoped per node to prevent backup collisions when running multiple autonomous trading nodes
- **`region`** - the storage region; Cloudflare R2 requires `"auto"`
- **`override_architecture`** - forces a specific binary architecture (`:x86_64`);
  useful when the autodetected architecture label doesn't match the download URL
- **`access_key_id`** / **`secret_access_key`** - S3 credentials

**Note:** We use Cloudflare R2 in this chapter, but Litestream works with any S3-compatible
storage. Backblaze B2, DigitalOcean Spaces, MinIO, and AWS S3 all work - adjust the
endpoint, bucket, and region accordingly. For AWS S3, you'd set `region` to your actual
region (e.g. `"us-east-1"`) instead of `"auto"`.

### Automatic restore on startup

Manual restores work, but they don't help when a server crashes at 3am and your
deployment pipeline spins up a fresh node. We want the application to heal itself:
if the database file is missing when the app boots, it should attempt a restore from
R2 automatically - before the Repo or any trading processes start.

We'll create a module that reads the same Litestream configuration we just defined,
checks if the database file exists, and if not, downloads the Litestream binary and
runs `restore -if-replica-exists`. The `-if-replica-exists` flag is important: on a
brand-new deployment where no backup has ever been made, it exits cleanly instead of
crashing - letting the app proceed to create a fresh database via migrations.

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/litestream/auto_restore.ex
defmodule Hedgehog.Litestream.AutoRestore do
  @moduledoc """
  Automatically restores the SQLite database from a Litestream backup
  if the database file is missing at startup.

  Called from application.ex before the supervision tree starts.
  """

  require Logger

  def maybe_restore do
    litestream_config = Application.get_env(:hedgehog, Hedgehog.Litestream, [])
    repo_config = Application.get_env(:hedgehog, Hedgehog.Repo, [])
    db_path = Keyword.get(repo_config, :database)

    cond do
      !Keyword.get(litestream_config, :enabled, false) ->
        :ok

      is_nil(db_path) ->
        Logger.warning("AutoRestore: no database path configured, skipping")
        :ok

      File.exists?(db_path) ->
        Logger.info("AutoRestore: database exists at #{db_path}, skipping restore")
        :ok

      true ->
        restore_from_backup(db_path, litestream_config)
    end
  end

  defp restore_from_backup(db_path, config) do
    Logger.info("AutoRestore: no database at #{db_path}, attempting restore from backup...")

    db_path |> Path.dirname() |> File.mkdir_p!()

    yaml_path =
      Path.join(System.tmp_dir!(), "litestream-restore-#{System.system_time(:millisecond)}.yml")

    bucket = Keyword.fetch!(config, :bucket)
    replica_path = Keyword.fetch!(config, :replica_path)
    endpoint = Keyword.fetch!(config, :endpoint)
    region = Keyword.get(config, :region, "auto")
    access_key_id = Keyword.fetch!(config, :access_key_id)
    secret_access_key = Keyword.fetch!(config, :secret_access_key)

    yaml_content = """
    access-key-id: "#{access_key_id}"
    secret-access-key: "#{secret_access_key}"

    dbs:
      - path: #{db_path}
        replica:
          url: s3://#{bucket}/#{replica_path}
          endpoint: #{endpoint}
          region: #{region}
    """

    File.write!(yaml_path, yaml_content)

    try do
      bin_path = ensure_binary(config)

      case System.cmd(bin_path, ["restore", "-config", yaml_path, "-if-replica-exists"],
             stderr_to_stdout: true
           ) do
        {output, 0} ->
          if File.exists?(db_path) do
            Logger.info("AutoRestore: database restored to #{db_path}")
          else
            Logger.info("AutoRestore: no replica found, starting with fresh database")
          end

          Logger.debug("Litestream output: #{output}")
          :ok

        {output, exit_code} ->
          Logger.warning(
            "AutoRestore: litestream exited with code #{exit_code}, " <>
              "starting with fresh database. Output: #{output}"
          )

          :ok
      end
    after
      File.rm(yaml_path)
    end
  end

  defp ensure_binary(config) do
    case Keyword.get(config, :bin_path) do
      nil ->
        download_dir = Path.join(System.tmp_dir!(), "litestream_downloads")
        bin_dir = Path.join(System.tmp_dir!(), "litestream_bin")
        File.mkdir_p!(download_dir)
        File.mkdir_p!(bin_dir)

        {:ok, path} =
          Litestream.Downloader.download_litestream(
            Litestream.Downloader.latest_version(),
            download_dir,
            bin_dir
          )

        path

      path ->
        path
    end
  end
end
```

A few things to note about this module. It never crashes the application - if the
restore fails for any reason(network issues, bad credentials, corrupt replica), it
logs a warning and returns `:ok`. The app will then start normally, create a fresh
database via migrations, and begin a new Litestream backup. You lose historical data,
but the node stays alive. The `-if-replica-exists` flag makes this safe for first-time
deployments: no replica in the bucket means Litestream exits cleanly and we get a
fresh database. The YAML config is written to a temp file and cleaned up in an `after`
block to avoid leaking credentials to the filesystem. And `ensure_binary/1` reuses the
same download mechanism that the Litestream GenServer uses, so the binary is cached
after the first download.

### Adding to the supervision tree

We want Litestream to start automatically in production and stay out of the way in
development. Update `application.ex` to conditionally include it:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/application.ex
  def start(_type, _args) do
    Hedgehog.Litestream.AutoRestore.maybe_restore()

    children =
      [
        ...
      ] ++ litestream_children()
    ...
  end

  defp litestream_children do
    config = Application.get_env(:hedgehog, Hedgehog.Litestream, [])

    if Keyword.get(config, :enabled, false) do
      opts = Keyword.drop(config, [:enabled])
      [{Litestream, opts}]
    else
      []
    end
  end
```

The `litestream_children/0` function returns an empty list when backups are disabled,
so the Litestream GenServer simply isn't started in dev or test. In production,
`maybe_restore()` runs first - restoring the database from R2 if the file is missing -
then the Repo starts against the restored(or freshly created) database, and the
Litestream GenServer begins replicating immediately.

When the GenServer starts for the first time, it downloads the Litestream binary
appropriate for your platform. Subsequent starts use the cached binary. Once running,
Litestream watches the database and continuously pushes changes to your S3 bucket.
You don't need to think about it again - it just runs in the background as a supervised
process, and the BEAM restarts it if it crashes.

### Setting up the storage bucket

Before deploying, you'll need an S3-compatible storage bucket. We'll use Cloudflare R2
because the setup is minimal and the free tier is generous(10 GB storage, zero egress
fees). Any S3-compatible service works - Backblaze B2, DigitalOcean Spaces, MinIO - but
R2 requires the least ceremony to get started.

Litestream speaks the S3 protocol, so it needs three things from your storage provider:
an **endpoint**(where to connect), an **Access Key ID** and a **Secret Access Key**(how
to authenticate). For Cloudflare R2, you'll need to:

1. Create a storage bucket - this is the container for your backup data. Pick a name
   that identifies the project(e.g. `hedgehog`).
2. Create an API token with read and write permissions scoped to that bucket. This
   generates an **Access Key ID** and **Secret Access Key** pair - save both immediately,
   as the secret is only displayed once.
3. Find your **Account ID** - a unique identifier for your Cloudflare account, visible
   on the R2 overview page. This becomes part of the endpoint URL.

The endpoint URL uses the format:

```
https://<account-id>.r2.cloudflarestorage.com
```

Set the three environment variables before starting your application:
```{r, engine = 'bash', eval = FALSE}
export LITESTREAM_REPLICA_URL="https://<your-account-id>.r2.cloudflarestorage.com"
export LITESTREAM_ACCESS_KEY_ID=<your-r2-access-key-id>
export LITESTREAM_SECRET_ACCESS_KEY=<your-r2-secret-access-key>
```

The bucket name(`hedgehog`) and backup path prefix are set in the Elixir
configuration rather than encoded in the URL. This keeps the environment variables
simple and makes it easy to change the backup path without touching production
environment settings.

### Restoring from a backup

With the auto-restore module in place, recovery from a complete server failure is
fully automated. Here's what happens when a fresh node boots with `DATABASE_PATH`
pointing to a non-existent file:

1. `application.ex` calls `AutoRestore.maybe_restore()` before starting the
   supervision tree
2. The module reads the Litestream config from `runtime.exs`, downloads the
   Litestream binary(if not already cached), and generates a temporary YAML config
3. It runs `litestream restore -if-replica-exists`, which pulls the latest snapshot
   from your R2 bucket into the configured `DATABASE_PATH`
4. The temp config(which contains S3 credentials) is cleaned up immediately
5. The supervision tree starts: the Repo opens the restored database, and the
   Litestream GenServer begins replicating new changes back to R2

If no replica exists yet(first-ever deployment), the restore exits cleanly
and the app creates a fresh database via migrations. If the restore fails for any
other reason(network issue, bad credentials), it logs a warning and continues -
you get a fresh database rather than a crashed node.

You can verify this works by deleting the database file and restarting:

```{r, engine = 'bash', eval = FALSE}
# In production (or a local test with Litestream enabled):
rm /data/hedgehog/hedgehog.db*
# restart the application
# watch the logs:
[info] AutoRestore: no database at /data/hedgehog/hedgehog.db, attempting restore from backup...
[info] AutoRestore: database restored to /data/hedgehog/hedgehog.db
```

#### Manual restore(fallback)

If you need to restore outside the application(for example, to inspect a backup
without starting the trading bot), you can use the Litestream CLI directly.

Install the Litestream binary that matches the version bundled by the Hex package(run
`mix deps` and look at the `litestream` entry to confirm the binary version). The
example below is for Linux x86_64 - adjust the archive name for your platform(e.g.
`litestream-0.5.8-darwin-amd64.tar.gz` for macOS, or `-linux-arm64` for ARM servers):

````{r, engine = 'bash', eval = FALSE}
curl -fsSL https://github.com/benbjohnson/litestream/releases/download/v0.5.8/litestream-0.5.8-linux-x86_64.tar.gz \
  | tar -xz -C /usr/local/bin
````

Then create a minimal config file for the restore. The CLI needs the same region and
endpoint settings that our Elixir config provides at runtime, and these can only be
specified through a config file:

```{r, engine = 'text', eval = FALSE}
# /tmp/litestream-restore.yml
access-key-id: ${LITESTREAM_ACCESS_KEY_ID}
secret-access-key: ${LITESTREAM_SECRET_ACCESS_KEY}

dbs:
  - path: /data/hedgehog/hedgehog.db
    replica:
      url: s3://hedgehog/backups/${HEDGEHOG_NODE_NAME}
      endpoint: ${LITESTREAM_REPLICA_URL}
      region: auto
```

**Note:** The bucket name `hedgehog` in the S3 URL must match the `bucket` value in your
`runtime.exs` Litestream configuration. If you changed the bucket name there, update
this restore config to match.

Litestream performs environment variable expansion in the config file, so the same
variables we set for our Elixir application work here too(plus the node name for
path isolation):

```{r, engine = 'bash', eval = FALSE}
export LITESTREAM_REPLICA_URL="https://<your-account-id>.r2.cloudflarestorage.com"
export LITESTREAM_ACCESS_KEY_ID=<your-r2-access-key-id>
export LITESTREAM_SECRET_ACCESS_KEY=<your-r2-secret-access-key>
export HEDGEHOG_NODE_NAME=<your-node-name>
litestream restore -config /tmp/litestream-restore.yml
```

Start your application normally after the restore. Litestream will pick up where it
left off and resume replication.

### Verifying replication locally (optional)

If you want to verify the setup works before deploying to production, Litestream also
supports replicating to a local file path. Temporarily update your `dev.exs`:

```{r, engine = 'elixir', eval = FALSE}
# /config/dev.exs (temporary - for testing only)
config :hedgehog, Hedgehog.Litestream,
  enabled: true,
  repo: Hedgehog.Repo,
  replica_url: "file://#{Path.expand("../backups/dev_node", Path.dirname(__ENV__.file))}",
  access_key_id: "unused",
  secret_access_key: "unused"
```

Start the application and check that a `backups/dev_node/` directory appears with replication
data inside. Note that we're using `replica_url` here instead of the decomposed
`endpoint`/`bucket`/`replica_path` from the production config - for `file://` destinations,
a single URL is all Litestream needs. Once you've confirmed it works, set `enabled: false`
back in `dev.exs` and remove the `backups/` directory.

### Checking replication status

You can check whether Litestream is running from an IEx session:

```{r, engine = 'bash', eval = FALSE}
iex -S mix phx.server
...
iex(1)> Litestream.status()
:running
```

If you need to temporarily stop replication (for example, during a maintenance window):

```{r, engine = 'bash', eval = FALSE}
iex(2)> Litestream.stop_litestream()
:ok
iex(3)> Litestream.status()
:down
iex(4)> Litestream.start_litestream()
:ok
```

## Final thoughts

**This was a short chapter, but the operational impact is significant.** We replaced
Postgres with SQLite, eliminating the Docker dependency entirely. We swapped
`ecto_enum`(Postgres-specific) for `Ecto.Enum`(database-agnostic) and updated status
values from strings to atoms across queries, update functions, and seeds. We rewrote all
migrations to use plain string columns instead of custom Postgres types, added continuous
backups via Litestream with per-node isolation, and wired up automatic restore so a
crashed node recovers its database before the Repo starts - all while preserving existing
tests and trading functionality without touching the business logic.

Notice what didn't change: the Repo API usage, the supervision tree, the trading logic.
What did change were the status value literals - `"on"` to `:on`, `"off"` to `:off` -
because `Ecto.Enum` is stricter than the old `ecto_enum` library about types.
That strictness is a feature: it catches type mismatches at query time instead of silently
inserting bad data.

The trade-off is real, though. SQLite doesn't support concurrent writes from multiple processes as
gracefully as Postgres does. If two GenServers try to write at the same time, one waits(up to `busy_timeout`
milliseconds) for the other to finish. For our use case - a trading bot running as a
single node writing its own local data - that's perfectly fine.
We're not building a multi-tenant web app; we're building independent trading nodes.

Two other limitations worth knowing if you apply this pattern elsewhere:
SQLite has no built-in access control - anyone with filesystem access to the `.db` file can read and write everything,
so security lives entirely at the OS level.

SQLite’s `ALTER TABLE` is more restrictive than Postgres -
you can add and drop columns, but you generally cannot modify
existing columns(such as changing types or adding constraints),
and many schema changes require recreating the table.

Future migrations may need to recreate tables rather than alter them.

And that's exactly where we're headed.

**What's next?**

We now have an application with no external infrastructure dependencies - no database
server to manage, no Docker container to run, just our Elixir app and a local
database file. Each node's backups land in their own S3 prefix, so multiple instances
can replicate simultaneously without stepping on each other. This is the perfect
building block for distribution. In the next chapter, we'll configure `libcluster`
for automatic cluster formation and start running multiple trading nodes side by side.
Each node will carry its own SQLite database, stream its own symbols, and execute its
own trades - while remaining connected via Erlang's distribution protocol for
coordination and monitoring.

[Note] Please remember to run `mix format` to keep things nice and tidy.

The source code for this chapter can be found in the book's source code repository
(branch:
[chapter_23](https://github.com/Cinderella-Man/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_23)).