# Clustering on the Open Internet

In the last chapter, we replaced Postgres with SQLite - each trading node now
carries its own database as a single file. We set up Litestream for continuous
backups and removed every external infrastructure dependency. The application
is fully self-contained.

But self-contained nodes sitting in isolation aren't very useful. We want to run
multiple trading nodes - potentially on different machines, in different data
centres, maybe even different continents - and still have a unified view of
what's happening across the fleet. Which symbols are being traded? How many
open positions exist? What's the total order volume? Right now, answering any
of these questions means SSH-ing into each machine and running queries by hand.

Erlang was built for exactly this problem. The BEAM has native support for
connecting nodes into a cluster, where any process on any node can communicate
with any process on any other node as if they were local. This is not an
add-on or a library feature - it's baked into the runtime. Two Erlang nodes
that know each other's names and share a secret cookie can form a cluster with
a single function call.

In this chapter, we'll build an OTP release, deploy it to plain virtual
machines(no containers), configure Erlang distribution for secure communication
over the open internet, wire up `libcluster` for automatic cluster formation,
and build a module that queries data across all connected nodes. By the end,
we'll have a fleet of independent trading bots that can see each other.

## Objectives
- building an OTP release for deployment
- deploying to virtual machines without containers
- understanding Erlang distribution: node names, cookies, and EPMD
- securing inter-node communication on the open internet
- configuring `libcluster` for automatic cluster formation
- querying data across a dynamic set of nodes

## Building an OTP release

An OTP release is a self-contained package of your application, its
dependencies, the Elixir runtime, and the Erlang VM - everything needed to
run without installing Elixir or Erlang on the target machine.
Since Elixir 1.9, the built-in `mix release` handles this
for us(older guides may reference a third-party tool called Distillery, which is no longer needed).

### Configuring the release in `mix.exs`

Before generating any files, let's configure the release itself. Every release
is tied to a version - this is the version in your `mix.exs`, and it matters
more than you might think. OTP uses this version to track what's running on
each node, and later(when we add hot code upgrades) to compute the
differences between the old and new code.

Make sure your `mix.exs` has an explicit version:

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
defmodule Hedgehog.MixProject do
  use Mix.Project

  def project do
    [
      app: :hedgehog,
      version: "0.1.0", # <= this version drives the release
      ...
      releases: releases()
    ]
  end

  ...

  defp releases do
    [
      hedgehog: [
        include_executables_for: [:unix],
        strip_beams: false, # <= added - keep debug info for upgrades
        steps: [:assemble, :tar] # <= added - produce a tarball
      ]
    ]
  end
end
```

Two things to note about this configuration. `strip_beams: false` preserves
debug information and module metadata in the compiled `.beam` files. Without
this, OTP cannot generate the instructions it needs when upgrading code on
a running node - we'll use this in the next chapter when we introduce hot code
reloading. The cost is slightly larger beam files(typically 20-40% bigger), but
for a trading bot this is negligible.

`steps: [:assemble, :tar]` tells `mix release` to produce a compressed tarball
alongside the normal release directory. This tarball is what we'll ship to
production servers. Deploying a tarball rather than syncing a directory is
important: it gives us a clean, versioned artifact that we can unpack on the
server without destroying the previous version's files. That matters for
rollbacks, and it matters even more when we add hot code upgrades.

### Generating release configuration files

Generate the template files that control how the release behaves:

```{r, engine = 'bash', eval = FALSE}
mix release.init
* creating rel/vm.args.eex
* creating rel/remote.vm.args.eex
* creating rel/env.sh.eex
* creating rel/env.bat.eex
```

This creates four files in the `rel/` directory. We care about two of them:
`env.sh.eex`(sets environment variables before the VM starts) and
`vm.args.eex`(passes flags directly to the Erlang VM).

### Configuring the node name and cookie

We'll configure the node name and cookie now to get the release built,
then explain the full mechanics of Erlang distribution after the deploy.
If you want the theory first, skip ahead to the "Understanding Erlang distribution"
section and come back here after.

For nodes to find and trust each other, every node needs two things: a unique
name and a shared secret called a cookie. We'll set both through environment
variables in `rel/env.sh.eex`:

```{r, engine = 'bash', eval = FALSE}
# /rel/env.sh.eex
export RELEASE_DISTRIBUTION=name
export RELEASE_NODE=${NODE_NAME:-hedgehog@127.0.0.1}
export RELEASE_COOKIE=${RELEASE_COOKIE:-$(cat /home/hedgehog/.erlang.cookie)}
```

A few things to note here. `RELEASE_DISTRIBUTION=name` tells the VM to use
fully qualified names(also called "long names") - this is required when nodes
run on different machines. The alternative, `sname`(short names), only works
within a single hostname or local network and cannot cross network boundaries.

`RELEASE_NODE` sets the node's identity. The format is `name@host`, where
`host` must be an IP address or a DNS name that other nodes can resolve. We
default to `hedgehog@127.0.0.1` for local development, but in production
you'll set `NODE_NAME` to something like `hedgehog@203.0.113.10`.

`RELEASE_COOKIE` is the shared secret. We read it from a file by default, but
you can also set it directly as an environment variable. The cookie must be
identical on every node that should form a cluster.

### Constraining the distribution port range

Erlang uses a daemon called EPMD(Erlang Port Mapper Daemon) to map node names
to ports. EPMD itself listens on port 4369, and each node picks a random high
port for its actual inter-node traffic. This is a problem for firewalls - you
can't whitelist a random port.

We fix this by telling the VM to use a specific port range. Add the following
to `rel/vm.args.eex`:

```{r, engine = 'text', eval = FALSE}
# /rel/vm.args.eex
-kernel inet_dist_listen_min 9100
-kernel inet_dist_listen_max 9155
```

This constrains the distribution port to a known range. For a single node per
machine, a range of one port(e.g. 9100-9100) is sufficient. A wider range
accommodates multiple nodes on the same machine during development or testing.

### Running database migrations in a release

In development, we run `mix ecto.migrate`. But a release doesn't have Mix -
it's a compiled artifact. Phoenix projects include a release module that
provides migration support. If your project doesn't have one, create it:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/release.ex
defmodule Hedgehog.Release do
  @app :hedgehog

  def migrate do
    load_app()

    for repo <- repos() do
      {:ok, _, _} =
        Ecto.Migrator.with_repo(repo, &Ecto.Migrator.run(&1, :up, all: true))
    end
  end

  def rollback(repo, version) do
    load_app()
    {:ok, _, _} = Ecto.Migrator.with_repo(repo, &Ecto.Migrator.run(&1, :down, to: version))
  end

  defp repos do
    Application.fetch_env!(@app, :ecto_repos)
  end

  defp load_app do
    Application.load(@app)
  end
end
```

Notice that `load_app/0` uses `Application.load/1` - not
`Application.ensure_all_started/1`. The difference is critical. `load/1` reads
the application's configuration without starting any processes.
`ensure_all_started/1` would boot the entire supervision tree - the Repo,
Litestream, Phoenix endpoint, trading processes - all before the migration
has run. That means Litestream would try to connect to S3, the Repo would try
to open a database that might not have the right schema yet, and you'd need
every production environment variable set just to run a migration.
`Ecto.Migrator.with_repo/2` handles starting just the Repo for us, which is
all a migration needs.

After deploying, run migrations with:

```{r, engine = 'bash', eval = FALSE}
bin/hedgehog eval "Hedgehog.Release.migrate()"
```

### Building the release

Build the release for production:

```{r, engine = 'bash', eval = FALSE}
MIX_ENV=prod mix release
...
* assembling hedgehog-0.1.0 on MIX_ENV=prod
* using config/runtime.exs to configure the release at runtime
* building hedgehog-0.1.0.tar.gz
...
Release created at _build/prod/rel/hedgehog
```

Because we configured `steps: [:assemble, :tar]`, the build produces both the
release directory *and* a tarball at
`_build/prod/rel/hedgehog/hedgehog-0.1.0.tar.gz`. The tarball is what we'll
ship to production. It contains everything - the ERTS runtime, your compiled
code, all dependencies - in a single file named after your application version.

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on cross-compilation:** The release must be built on the same OS and
architecture as the target machine. If you develop on macOS but deploy to
Linux, you need to build the release on a Linux machine(or in a Linux VM/CI pipeline).
You cannot build on `x86_64-apple-darwin` and deploy to `x86_64-unknown-linux-gnu` -
the ERTS binary and any native code(including the SQLite NIF used by `ecto_sqlite3`) won't run.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

## Deploying to a virtual machine

We're deploying to plain VMs - no Docker, no Kubernetes, no orchestrator.
Just Linux machines with an SSH connection. This is the simplest production
setup and it maps directly to our architecture: one machine, one node, one
database.

### What you need on the target machine

The target machine needs surprisingly little:

- A Linux distribution(Ubuntu, Debian, etc.)
- A user account to run the application(don't run as root)
- The `libncurses` and `libssl` libraries(required by ERTS)
- Directories for the application and database

Set up the target machine:

```{r, engine = 'bash', eval = FALSE}
# On the target machine
sudo apt-get update && sudo apt-get install -y libncurses5 libssl-dev
sudo useradd -m -s /bin/bash hedgehog
sudo mkdir -p /opt/hedgehog /data/hedgehog
sudo chown hedgehog:hedgehog /opt/hedgehog /data/hedgehog
```

### Deploying the release

Copy the tarball to the target machine and unpack it:

```{r, engine = 'bash', eval = FALSE}
# From the build machine
scp _build/prod/rel/hedgehog/hedgehog-0.1.0.tar.gz \
  hedgehog@your-server:/opt/hedgehog/
```

```{r, engine = 'bash', eval = FALSE}
# On the target machine, as the hedgehog user
cd /opt/hedgehog
tar -xzf hedgehog-0.1.0.tar.gz
rm hedgehog-0.1.0.tar.gz
```

After unpacking, `/opt/hedgehog` contains the standard OTP release layout:

```
/opt/hedgehog/
├── bin/
│   └── hedgehog          # CLI entry point
├── erts-<version>/       # Erlang runtime
├── lib/
│   ├── hedgehog-0.1.0/   # your compiled code (versioned!)
│   ├── ecto-3.x.x/       # dependencies (versioned)
│   └── ...
└── releases/
    ├── 0.1.0/            # release metadata for this version
    │   ├── sys.config
    │   └── vm.args
    ├── RELEASES          # tracks installed versions
    └── start_erl.data
```

Notice how everything is versioned - `lib/hedgehog-0.1.0/`,
`releases/0.1.0/`. This is the OTP convention and it matters. When we deploy
version `0.2.0` later, the new directories sit alongside the old ones rather
than replacing them. The running node can serve the old code while the new
code is being installed - which is exactly what makes hot code upgrades
possible.

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Why not rsync?** You might be tempted to use `rsync -avz` to push the
release directory directly. This works for a first deploy but creates problems
later: rsync overwrites existing files, which means the old version's beam
files are destroyed. If you ever need to rollback, or if you want to upgrade
code on a running node without restarting it, you need both the old and new
versions on disk simultaneously. Deploying via tarball preserves this property
naturally - each version is a clean, self-contained snapshot.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

### Environment variables

Create an environment file with all the variables our application needs.
This includes the database path from chapter 23, the Litestream credentials,
and the new distribution settings:

```{r, engine = 'bash', eval = FALSE}
# /etc/hedgehog/env
DATABASE_PATH=/data/hedgehog/hedgehog.db
NODE_NAME=hedgehog@203.0.113.10
RELEASE_COOKIE=a_very_long_random_string_that_all_nodes_share

# Phoenix
SECRET_KEY_BASE=a_very_long_random_string_for_phoenix
PORT=4000

# Litestream (from chapter 23)
LITESTREAM_REPLICA_URL=s3://hedgehog-backups/node1-db?endpoint=<account-id>.r2.cloudflarestorage.com
LITESTREAM_ACCESS_KEY_ID=<your-access-key-id>
LITESTREAM_SECRET_ACCESS_KEY=<your-secret-access-key>

# Binance API
BINANCE_API_KEY=<your-api-key>
BINANCE_API_SECRET=<your-api-secret>
```

Set the file permissions so only the `hedgehog` user can read it:

```{r, engine = 'bash', eval = FALSE}
sudo chmod 600 /etc/hedgehog/env
sudo chown hedgehog:hedgehog /etc/hedgehog/env
```

### Creating a systemd service

We want the application to start automatically on boot, restart on crashes,
and log to the system journal. A systemd service handles all of this:

```{r, engine = 'text', eval = FALSE}
# /etc/systemd/system/hedgehog.service
[Unit]
Description=Hedgehog Trading Bot
After=network.target

[Service]
Type=exec
User=hedgehog
Group=hedgehog
EnvironmentFile=/etc/hedgehog/env
WorkingDirectory=/opt/hedgehog
ExecStartPre=/opt/hedgehog/bin/hedgehog eval "Hedgehog.Release.migrate()"
ExecStart=/opt/hedgehog/bin/hedgehog start
ExecStop=/opt/hedgehog/bin/hedgehog stop
Restart=on-failure
RestartSec=5
LimitNOFILE=65536

[Install]
WantedBy=multi-user.target
```

The `ExecStartPre` line runs migrations before the application starts -
this ensures the database schema is up to date on every deploy. Enable and
start the service:

```{r, engine = 'bash', eval = FALSE}
sudo systemctl daemon-reload
sudo systemctl enable hedgehog
sudo systemctl start hedgehog
```

Check that it's running:

```{r, engine = 'bash', eval = FALSE}
sudo systemctl status hedgehog
● hedgehog.service - Hedgehog Trading Bot
     Active: active (running)
...
sudo journalctl -u hedgehog -f
...
```

### Connecting to a running node

One of the best things about OTP releases is remote shell access. You can
attach to a running production node and inspect it interactively:

```{r, engine = 'bash', eval = FALSE}
# On the target machine, as the hedgehog user
/opt/hedgehog/bin/hedgehog remote
Erlang/OTP ...
Interactive Elixir ...
iex(hedgehog@203.0.113.10)1> Node.self()
:"hedgehog@203.0.113.10"
iex(hedgehog@203.0.113.10)2> Node.get_cookie()
:a_very_long_random_string_that_all_nodes_share
```

You're now inside the running VM. You can inspect processes, query the
database, check Litestream status - anything you could do in a local IEx
session. Press `Ctrl+C` twice to disconnect without stopping the node.

### Deploying an update

When you're ready to deploy a new version, bump the version in `mix.exs`,
rebuild, and ship the new tarball:

```{r, engine = 'bash', eval = FALSE}
# After updating mix.exs to version "0.2.0" and rebuilding:
scp _build/prod/rel/hedgehog/hedgehog-0.2.0.tar.gz \
  hedgehog@your-server:/opt/hedgehog/
```

```{r, engine = 'bash', eval = FALSE}
# On the server - stop, unpack, start
sudo systemctl stop hedgehog
cd /opt/hedgehog
tar -xzf hedgehog-0.2.0.tar.gz
rm hedgehog-0.2.0.tar.gz
sudo systemctl start hedgehog
```

After unpacking, `lib/` contains both `hedgehog-0.1.0/` and
`hedgehog-0.2.0/`, and `releases/` contains both `0.1.0/` and `0.2.0/`. The
node starts with the latest version. The old version's files remain on disk -
this is intentional and will become essential in the next chapter when we
upgrade without stopping.

For now, this stop-unpack-start cycle is our deployment workflow. It's simple,
predictable, and the brief downtime(typically under 5 seconds) is acceptable
for an initial deployment. In the next chapter, we'll eliminate that downtime
entirely.

## Understanding Erlang distribution

Before we configure automatic clustering, it's worth understanding what
happens under the hood when two Erlang nodes connect. This knowledge will
save you hours of debugging when things don't work.

### Node names

Every distributed Erlang node has a name in the format `name@host`. The `name`
part is arbitrary - we use `hedgehog`. The `host` part must be something the
other nodes can resolve: an IP address(`hedgehog@203.0.113.10`) or a DNS
name(`hedgehog@node1.example.com`).

There are two naming modes:

- **Short names**(`-sname`): The host part is just a hostname without a domain.
  Nodes using short names can only connect to other nodes on the same machine
  or the same local network. Fine for development, not for production across
  the internet.

- **Long names**(`-name`): The host part is a fully qualified domain name or
  IP address. Required for nodes that need to connect across networks. This
  is what we configured with `RELEASE_DISTRIBUTION=name`.

A node using short names cannot connect to a node using long names - they must
all agree on the naming mode.

### The cookie

The cookie is a shared secret that acts as a first line of authentication.
When node A tries to connect to node B, both nodes compare cookies. If they
don't match, the connection is refused. That's it - there's no encryption,
no challenge-response, no certificate exchange. Just a string comparison.

This has an important security implication: anyone who knows your cookie and
can reach your node's distribution port has full access to your VM. They can
execute arbitrary code, read your data, and shut down your application. We'll
address this in the security section.

Generate a strong cookie:

```{r, engine = 'bash', eval = FALSE}
elixir -e "IO.puts(Base.encode32(:crypto.strong_rand_bytes(40)))"
W3VFHXQW2Q3EZBSGK7GWCPDCW36LVWBXD2NOYF54DXQLQYVW7YVA====
```

Use this value as `RELEASE_COOKIE` on every node. Don't use short, guessable
strings like `hedgehog` or `secret` - treat the cookie like a password.

### EPMD

EPMD(Erlang Port Mapper Daemon) is a small service that runs on every machine
hosting Erlang nodes. It listens on TCP port 4369 and maintains a registry of
node names to distribution ports. When node A wants to connect to
`hedgehog@203.0.113.10`, it first contacts EPMD on `203.0.113.10:4369` and
asks "what port is `hedgehog` listening on?". EPMD responds with the port
number, and node A connects directly.

EPMD starts automatically when you start a distributed Erlang node. You don't
need to manage it separately, but you need to know it exists because:

- Port 4369 must be reachable between nodes(or at least between the connecting
  node and the target's EPMD)
- The distribution port(which we constrained to 9100-9155) must also be
  reachable
- EPMD itself has no authentication - anyone can query it to discover which
  nodes are running on a machine

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on EPMD alternatives:** Erlang supports running without EPMD by using a custom discovery module.
The `-epmd_module` VM flag has supported this for years, though it was deprecated in OTP 27.
The replacement is the `epmd_module` kernel application parameter, available since OTP 27.2.
For our setup with `libcluster` handling discovery, EPMD is still the simplest path,
but eliminating it reduces your attack surface by one open port.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

## Securing the cluster

Here's the uncomfortable truth about Erlang distribution: it was designed for
trusted networks. The cookie provides only a minimal gate - once past it, a
connected node can execute arbitrary Erlang code on any other node in the
cluster. There is no granular permission model. A connected node has the same
power as a local IEx session.

This means running an Erlang cluster on the open internet without additional
protection is asking for trouble. Let's fix that.

### The threat model

For our trading bot, we need to protect against:

- **Unauthorized connections:** Someone discovering our EPMD port and
  brute-forcing(or sniffing) the cookie
- **Eavesdropping:** Inter-node traffic(including database queries and trade
  data) is sent in cleartext by default
- **Port scanning:** Exposed EPMD and distribution ports advertising our
  cluster's existence

### Layer 1: Firewall rules

The first and most important defence is a firewall. Only allow distribution
traffic from known node IPs:

```{r, engine = 'bash', eval = FALSE}
# Allow EPMD and distribution ports only from known nodes
sudo ufw allow from 203.0.113.11 to any port 4369
sudo ufw allow from 203.0.113.11 to any port 9100:9155 proto tcp
sudo ufw allow from 203.0.113.12 to any port 4369
sudo ufw allow from 203.0.113.12 to any port 9100:9155 proto tcp
# Deny everything else on these ports (default deny)
sudo ufw enable
```

This is the single most effective measure. Even if someone obtains your cookie,
they can't connect if the firewall blocks their IP.

### Layer 2: A strong cookie

Use the long, random cookie we generated earlier. Avoid anything guessable.
The cookie is transmitted during the initial handshake but not used for ongoing
encryption - it's more like a shared password than a session key.

### Layer 3: VPN or WireGuard(recommended for production)

For a production deployment, the gold standard is to run all inter-node traffic
over a VPN. WireGuard is particularly well-suited for this: it's fast, simple
to configure, and adds encryption with minimal overhead.

With WireGuard, each node gets a private IP(e.g. `10.0.0.1`, `10.0.0.2`) and
all distribution traffic flows through the encrypted tunnel. Your node names
use the WireGuard IPs(`hedgehog@10.0.0.1`), the distribution ports are never
exposed to the public internet, and the cookie becomes a secondary safeguard
rather than your only protection.

Setting up WireGuard is outside the scope of this chapter, but the
configuration is straightforward - typically under 20 lines per node. The
WireGuard documentation and the `wg-quick` tool make it approachable even if
you haven't used a VPN before.

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on TLS distribution:** Erlang also supports encrypting the distribution
protocol itself using TLS, via the `-proto_dist inet_tls` flag. This provides
encryption and mutual certificate authentication without a VPN. However, it
requires managing TLS certificates on every node, configuring a custom
`ssl_dist_opt` file, and is considerably more complex to set up and maintain
than a WireGuard tunnel. If you're already operating a PKI(Public Key
Infrastructure) or are deploying in an environment where VPNs are not an
option, TLS distribution is a solid choice. See the Erlang documentation on
"Using TLS for Erlang Distribution" for the full setup procedure.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

### Summary of ports to open

For quick reference, here are the ports each node needs:

| Port | Protocol | Purpose |
|------|----------|---------|
| 4369 | TCP | EPMD(node discovery) |
| 9100-9155 | TCP | Erlang distribution(inter-node traffic) |
| 4000 | TCP | Phoenix web server(if using the web UI) |

Open these ports only to known node IPs(for 4369 and 9100-9155) or to
your management network. Port 4000 can be exposed more broadly if you want
external access to the web interface.

## Configuring `libcluster`

We could connect nodes manually using `Node.connect/1`, but that's tedious
and fragile. If a node restarts, all its connections are lost and someone
(or something) has to reconnect it. `libcluster` automates this - it
continuously attempts to form and heal a cluster based on a configured
strategy.

### Adding the dependency

```{r, engine = 'elixir', eval = FALSE}
# /mix.exs
  defp deps do
    [
      ...
      {:libcluster, "~> 3.5"}, # <= added
      ...
    ]
  end
```

```{r, engine = 'bash', eval = FALSE}
mix deps.get
...
```

### Choosing a strategy

`libcluster` ships with several clustering strategies. For VMs with known
IP addresses, `Cluster.Strategy.Epmd` is the right choice - you give it
a list of node names and it connects to them on startup. Other strategies
exist for dynamic environments(Kubernetes, multicast gossip, DNS), but for
a fixed set of VMs, a simple list is the most reliable and debuggable option.

### Configuration per environment

In development, we disable clustering or use a local gossip strategy for
testing. In production, we use the EPMD strategy with node names from
environment variables.

```{r, engine = 'elixir', eval = FALSE}
# /config/dev.exs
config :libcluster,
  topologies: []
```

```{r, engine = 'elixir', eval = FALSE}
# /config/test.exs
config :libcluster,
  topologies: []
```

```{r, engine = 'elixir', eval = FALSE}
# /config/integration.exs
config :libcluster,
  topologies: []
```

For production, we read the list of peer nodes from an environment variable.
Add the following inside the `if config_env() == :prod do` block in
`runtime.exs`:

```{r, engine = 'elixir', eval = FALSE}
# /config/runtime.exs (inside the :prod block)
  cluster_nodes =
    case System.get_env("CLUSTER_NODES") do
      nil ->
        []

      nodes_string ->
        nodes_string
        |> String.split(",", trim: true)
        |> Enum.map(&String.trim/1)
        |> Enum.map(&String.to_atom/1)
    end

  config :libcluster,
    topologies: [
      hedgehog: [
        strategy: Cluster.Strategy.Epmd,
        config: [
          hosts: cluster_nodes
        ]
      ]
    ]
```

The `CLUSTER_NODES` environment variable is a comma-separated list of node
names. On each node, you list all the *other* nodes it should connect to:

```{r, engine = 'bash', eval = FALSE}
# On node 203.0.113.10
export CLUSTER_NODES="hedgehog@203.0.113.11,hedgehog@203.0.113.12"

# On node 203.0.113.11
export CLUSTER_NODES="hedgehog@203.0.113.10,hedgehog@203.0.113.12"

# On node 203.0.113.12
export CLUSTER_NODES="hedgehog@203.0.113.10,hedgehog@203.0.113.11"
```

Add `CLUSTER_NODES` to each node's `/etc/hedgehog/env` file.

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on Erlang's mesh topology:** Erlang clusters are fully connected by
default - every node connects to every other node. When node A connects to
node B, and node B is already connected to node C, node A will automatically
connect to node C as well. This means you technically only need to list *one*
existing node in `CLUSTER_NODES` - once connected, the new node discovers and
connects to all others. However, listing all nodes is more resilient: if the
single listed node is down when you start, the new node won't join the cluster
at all.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

### Adding to the supervision tree

`libcluster` provides a `Cluster.Supervisor` that manages the connection
strategy. Add it to your supervision tree in `application.ex`, before the
trading supervisors but after the Repo:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/application.ex
  def start(_type, _args) do
    topologies = Application.get_env(:libcluster, :topologies, [])

    children =
      [
        Hedgehog.Repo,
        {Cluster.Supervisor, [topologies, [name: Hedgehog.ClusterSupervisor]]}, # <= added
        ...
      ] ++ litestream_children()
    ...
  end
```

The `Cluster.Supervisor` reads the topology configuration and immediately
begins connecting to the listed nodes. If a connection fails, it retries
automatically. If a connected node goes down, `libcluster` detects the
disconnection(via Erlang's built-in node monitoring) and attempts to reconnect.

## Testing the cluster locally

Before deploying to production, let's verify that clustering works by running
multiple nodes on our development machine.

### Starting multiple nodes

Open three terminal windows and start a node in each:

```{r, engine = 'bash', eval = FALSE}
# Terminal 1
iex --name hedgehog1@127.0.0.1 --cookie dev_secret -S mix phx.server
```

```{r, engine = 'bash', eval = FALSE}
# Terminal 2 (use a different port for Phoenix)
PORT=4001 iex --name hedgehog2@127.0.0.1 --cookie dev_secret -S mix phx.server
```

```{r, engine = 'bash', eval = FALSE}
# Terminal 3
PORT=4002 iex --name hedgehog3@127.0.0.1 --cookie dev_secret -S mix phx.server
```

### Connecting nodes manually

Since we disabled `libcluster` in dev, we can connect nodes manually to test
the fundamentals:

```{r, engine = 'bash', eval = FALSE}
# In Terminal 1
iex(hedgehog1@127.0.0.1)1> Node.connect(:"hedgehog2@127.0.0.1")
true
iex(hedgehog1@127.0.0.1)2> Node.connect(:"hedgehog3@127.0.0.1")
true
iex(hedgehog1@127.0.0.1)3> Node.list()
[:"hedgehog2@127.0.0.1", :"hedgehog3@127.0.0.1"]
```

Check from Terminal 2 - it should see both other nodes thanks to Erlang's
automatic mesh formation:

```{r, engine = 'bash', eval = FALSE}
# In Terminal 2
iex(hedgehog2@127.0.0.1)1> Node.list()
[:"hedgehog1@127.0.0.1", :"hedgehog3@127.0.0.1"]
```

The cluster is live. Every node can see every other node.

### Testing with libcluster locally

If you want to test `libcluster` itself, temporarily update `dev.exs`:

```{r, engine = 'elixir', eval = FALSE}
# /config/dev.exs (temporary - for testing only)
config :libcluster,
  topologies: [
    hedgehog: [
      strategy: Cluster.Strategy.Epmd,
      config: [
        hosts: [:"hedgehog1@127.0.0.1", :"hedgehog2@127.0.0.1", :"hedgehog3@127.0.0.1"]
      ]
    ]
  ]
```

Start the three nodes as before. This time they should connect automatically
without any manual `Node.connect/1` calls. Check the logs for messages from
`libcluster` confirming the connections, then reset `dev.exs` back to
`topologies: []` when you're done.

## Querying data across the cluster

Now for the payoff. We have multiple trading nodes, each with its own SQLite
database, connected in a cluster. How do we get a unified view of what's
happening across all of them?

The key insight is that each node already has everything it needs to answer
questions about itself - it has a `Hedgehog.Repo`, schemas, and query
functions. All we need is a way to call those functions remotely and merge
the results.

### Introducing `:erpc`

Erlang provides the `:erpc` module(introduced in OTP 23) for executing
functions on remote nodes. It improves on the older `:rpc` module with better
error handling and the ability to distinguish between returned values, raised
exceptions, and communication failures.

The function we care about is `:erpc.multicall/5` - it executes a function on
multiple nodes in parallel and collects the results:

```{r, engine = 'bash', eval = FALSE}
iex(hedgehog1@127.0.0.1)1> :erpc.multicall(Node.list(), IO, :puts, ["hello from remote"])
hello from remote
hello from remote
[{:ok, :ok}, {:ok, :ok}]
```

Each result is tagged with `{:ok, value}` on success or `{:error, reason}` on
failure. This makes it easy to handle partial results when some nodes are
unreachable.

### Building a cluster query module

Let's build a `Hedgehog.Cluster` module that provides a clean API for
cross-node queries:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/cluster.ex
defmodule Hedgehog.Cluster do
  @moduledoc """
  Query and aggregate data across all connected cluster nodes.
  """

  import Ecto.Query

  alias Hedgehog.Repo
  alias Hedgehog.Strategy.Naive.Settings, as: NaiveSettings

  @default_timeout 5_000

  @doc """
  Returns a list of all connected nodes, including the current one.
  """
  def all_nodes do
    [Node.self() | Node.list()]
  end

  @doc """
  Executes a function on all nodes (including this one) and collects results.
  Failed nodes are excluded from the result.
  """
  def multicall(module, function, args, timeout \\ @default_timeout) do
    nodes = all_nodes()
    results = :erpc.multicall(nodes, module, function, args, timeout)

    nodes
    |> Enum.zip(results)
    |> Enum.flat_map(fn
      {_node, {:ok, value}} when is_list(value) -> value
      {_node, {:ok, value}} -> [value]
      {node, {:error, reason}} ->
        require Logger
        Logger.warning("Cluster call to #{node} failed: #{inspect(reason)}")
        []
    end)
  end

  @doc """
  Executes a function on all nodes and returns results tagged with the node name.
  """
  def multicall_tagged(module, function, args, timeout \\ @default_timeout) do
    nodes = all_nodes()
    results = :erpc.multicall(nodes, module, function, args, timeout)

    nodes
    |> Enum.zip(results)
    |> Enum.map(fn
      {node, {:ok, value}} -> {node, {:ok, value}}
      {node, {:error, reason}} -> {node, {:error, reason}}
    end)
  end
end
```

### Querying orders across the cluster

Now we can build specific query functions. Let's add a function that returns
the trading status of all symbols across every node:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/cluster.ex (add to the existing module)

  @doc """
  Returns all naive strategy settings with status :on across all nodes.
  """
  def active_trading_symbols do
    multicall(__MODULE__, :local_active_symbols, [])
  end

  @doc false
  def local_active_symbols do
    Repo.all(
      from(s in NaiveSettings,
        where: s.status == :on,
        select: %{symbol: s.symbol, node: ^to_string(Node.self())}
      )
    )
  end
```

The pattern is simple: define a `local_*` function that queries the local
database, then call it across all nodes via `multicall/4`. Each node executes
the query against its own SQLite database and returns the result. The calling
node merges everything into a single list.

Try it from any node in the cluster:

```{r, engine = 'bash', eval = FALSE}
iex(hedgehog1@127.0.0.1)1> Hedgehog.Cluster.active_trading_symbols()
[
  %{symbol: "XRPUSDT", node: "hedgehog1@127.0.0.1"},
  %{symbol: "BTCUSDT", node: "hedgehog2@127.0.0.1"},
  %{symbol: "ETHUSDT", node: "hedgehog2@127.0.0.1"},
  %{symbol: "DOGEUSDT", node: "hedgehog3@127.0.0.1"}
]
```

One function call, one result, from every node in the cluster. Each node
handles its own symbols but you can see the full picture from anywhere.

### Adding a cluster status summary

Let's add one more practical function - a high-level summary of the cluster
state that's useful for monitoring:

```{r, engine = 'elixir', eval = FALSE}
# /lib/hedgehog/cluster.ex (add to the existing module)

  @doc """
  Returns a summary of each node's status: name, connected state,
  and number of active trading symbols.
  """
  def status do
    results = multicall_tagged(__MODULE__, :local_status, [])

    Enum.map(results, fn
      {node, {:ok, status}} ->
        Map.put(status, :node, node)

      {node, {:error, reason}} ->
        %{node: node, status: :unreachable, reason: reason}
    end)
  end

  @doc false
  def local_status do
    active_count =
      Repo.aggregate(
        from(s in NaiveSettings, where: s.status == :on),
        :count
      )

    %{
      status: :running,
      version: Application.spec(:hedgehog, :vsn) |> to_string(),
      active_symbols: active_count,
      uptime_seconds: :erlang.statistics(:wall_clock) |> elem(0) |> div(1000)
    }
  end
```

```{r, engine = 'bash', eval = FALSE}
iex(hedgehog1@127.0.0.1)1> Hedgehog.Cluster.status()
[
  %{node: :"hedgehog1@127.0.0.1", status: :running, version: "0.1.0", active_symbols: 1, uptime_seconds: 3842},
  %{node: :"hedgehog2@127.0.0.1", status: :running, version: "0.1.0", active_symbols: 2, uptime_seconds: 7201},
  %{node: :"hedgehog3@127.0.0.1", status: :running, version: "0.1.0", active_symbols: 1, uptime_seconds: 1923}
]
```

Notice the `version` field - we pull it from `Application.spec/2`, which reads
the version compiled into the release. This becomes particularly useful during
rolling deploys: if you've updated some nodes but not others, `status/0` will
show you exactly which version each node is running. In the next chapter, it
will tell you whether a hot code upgrade has been applied.

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on `:erpc.multicall` and timeouts:** The timeout in `:erpc.multicall/5`
is a *per-node* timeout, not a total timeout. If you set it to 5000ms and have
three nodes, the worst case is still ~5 seconds(because calls run in parallel),
not 15 seconds. However, if a node is unreachable and TCP hasn't timed out yet,
the call to that node may block for the full timeout duration. Keep this in mind
when choosing your timeout value - too short and you'll get false negatives on
slow networks; too long and a single dead node will make your status check feel
sluggish.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

```{=latex}
\begin{sidequest}
```
```{=html}
<div class="sidequest">
```

**Note on code consistency:** When using `:erpc` to call functions on remote
nodes, the code on the remote node must have the module and function you're
calling. This is obvious when all nodes run the same release, but it becomes
important during rolling deployments - if node A has been updated with a new
version of `Hedgehog.Cluster` that adds a function, but node B is still
running the old version, the `:erpc` call to node B will fail with an
`UndefinedFunctionError`. Keep this in mind when deploying updates across a
live cluster. We'll look at strategies for managing this in the next chapter
when we discuss hot code reloading.

```{=html}
</div>
```
```{=latex}
\end{sidequest}
```

## Deploying a second node

Let's walk through adding a second node to the cluster. Assume the first node
is already running on `203.0.113.10`.

### Provision the second machine

Repeat the target machine setup on `203.0.113.11`:

```{r, engine = 'bash', eval = FALSE}
# On 203.0.113.11
sudo apt-get update && sudo apt-get install -y libncurses5 libssl-dev
sudo useradd -m -s /bin/bash hedgehog
sudo mkdir -p /opt/hedgehog /data/hedgehog
sudo chown hedgehog:hedgehog /opt/hedgehog /data/hedgehog
```

### Deploy the release

```{r, engine = 'bash', eval = FALSE}
# From the build machine
scp _build/prod/rel/hedgehog/hedgehog-0.1.0.tar.gz \
  hedgehog@203.0.113.11:/opt/hedgehog/
```

```{r, engine = 'bash', eval = FALSE}
# On 203.0.113.11, as the hedgehog user
cd /opt/hedgehog
tar -xzf hedgehog-0.1.0.tar.gz
rm hedgehog-0.1.0.tar.gz
```

### Configure the environment

Create `/etc/hedgehog/env` on the second machine. The `RELEASE_COOKIE` must
match the first node. The `NODE_NAME` must be different. `CLUSTER_NODES` lists
the first node:

```{r, engine = 'bash', eval = FALSE}
# /etc/hedgehog/env on 203.0.113.11
DATABASE_PATH=/data/hedgehog/hedgehog.db
NODE_NAME=hedgehog@203.0.113.11
RELEASE_COOKIE=a_very_long_random_string_that_all_nodes_share
CLUSTER_NODES=hedgehog@203.0.113.10

SECRET_KEY_BASE=a_very_long_random_string_for_phoenix
PORT=4000

LITESTREAM_REPLICA_URL=s3://hedgehog-backups/node2-db?endpoint=<account-id>.r2.cloudflarestorage.com
LITESTREAM_ACCESS_KEY_ID=<your-access-key-id>
LITESTREAM_SECRET_ACCESS_KEY=<your-secret-access-key>

BINANCE_API_KEY=<your-api-key>
BINANCE_API_SECRET=<your-api-secret>
```

**Note:** Each node should use a different path prefix in `LITESTREAM_REPLICA_URL`
(e.g. `node1-db`, `node2-db`) so their backups don't overwrite each other.
Alternatively, use separate buckets per node.

### Open firewall ports

On both machines, allow distribution traffic from the other:

```{r, engine = 'bash', eval = FALSE}
# On 203.0.113.10
sudo ufw allow from 203.0.113.11 to any port 4369
sudo ufw allow from 203.0.113.11 to any port 9100:9155 proto tcp

# On 203.0.113.11
sudo ufw allow from 203.0.113.10 to any port 4369
sudo ufw allow from 203.0.113.10 to any port 9100:9155 proto tcp
```

### Update the first node's configuration

Add the new node to the first node's `CLUSTER_NODES`:

```{r, engine = 'bash', eval = FALSE}
# /etc/hedgehog/env on 203.0.113.10 - update this line:
CLUSTER_NODES=hedgehog@203.0.113.11
```

Restart the first node to pick up the change:

```{r, engine = 'bash', eval = FALSE}
sudo systemctl restart hedgehog
```

### Start the second node

```{r, engine = 'bash', eval = FALSE}
# On 203.0.113.11
sudo cp /etc/systemd/system/hedgehog.service /etc/systemd/system/hedgehog.service
sudo systemctl daemon-reload
sudo systemctl enable hedgehog
sudo systemctl start hedgehog
```

### Verify the cluster

Connect to either node and check:

```{r, engine = 'bash', eval = FALSE}
/opt/hedgehog/bin/hedgehog remote
iex(hedgehog@203.0.113.10)1> Node.list()
[:"hedgehog@203.0.113.11"]
iex(hedgehog@203.0.113.10)2> Hedgehog.Cluster.status()
[
  %{node: :"hedgehog@203.0.113.10", status: :running, version: "0.1.0", active_symbols: 1, uptime_seconds: 42},
  %{node: :"hedgehog@203.0.113.11", status: :running, version: "0.1.0", active_symbols: 0, uptime_seconds: 15}
]
```

The cluster is live. Two independent trading nodes, each with its own database,
connected and queryable from either side.

## Troubleshooting

Clustering problems are almost always one of four things. Here's how to
diagnose each:

**Nodes can't connect at all:**
Check that EPMD is running on both machines(`epmd -names`), that port 4369 is
reachable from the other node(`nc -zv 203.0.113.11 4369`), and that the
distribution port range(9100-9155) is open. Also verify that both nodes use
the same naming mode(`-name`, not `-sname`).

**Cookie mismatch:**
If the cookie doesn't match, you'll see a log message like
`connection attempt from disallowed node`. Verify cookies match with
`Node.get_cookie()` on each node.

**DNS resolution failures:**
If you use DNS names instead of IP addresses in your node names, every node
must be able to resolve every other node's hostname. Test with `nslookup` or
`dig` from each machine.

**Nodes connect then immediately disconnect:**
This usually means a network timeout or a firewall rule that allows the initial
TCP handshake but drops the connection shortly after. Check for stateful firewall
rules and ensure the distribution port(not just EPMD) is whitelisted.

## Final thoughts

**We now have a distributed trading system.** Multiple independent nodes, each
running its own trading strategies against its own local SQLite database,
connected via Erlang's native distribution protocol. Any node can query any
other node's data with a single function call. Nodes can join and leave the
cluster dynamically - `libcluster` handles reconnection automatically.

The architecture is deliberately simple. There's no central coordinator, no
message broker, no shared database. Each node is self-sufficient: it can trade
on its own even if every other node in the cluster goes down. The cluster adds
visibility and coordination, not dependency.

Notice what we didn't do: we didn't build consensus protocols, distributed
transactions, or shared state. Each node owns its own data. The `Hedgehog.Cluster`
module is a read-only aggregation layer - it queries remote nodes but never
writes to them. This keeps the system simple and avoids the hardest problems
in distributed computing. If you need a node to change its behavior, you
connect to it directly and tell it what to do.

The security setup is layered: firewall rules block unauthorized access, a
strong random cookie prevents accidental connections, and a VPN(WireGuard)
encrypts everything in transit. You don't need all three layers, but each one
reduces your attack surface.

We also made several choices in this chapter that might have seemed like minor
details but were actually laying groundwork. The `strip_beams: false` option
preserves the metadata OTP needs to compute code differences. The tarball-based
deployment keeps old version directories intact alongside new ones. The
`version` field in `Cluster.status/0` tells you exactly what each node is
running. These pieces all come together in the next chapter.

One thing we don't handle is network partitions. If the network splits and
half the nodes can't see the other half, both halves keep running independently.
Since `Hedgehog.Cluster` is read-only, the worst case is an incomplete status view.
If we were coordinating writes across nodes, we'd need a consensus protocol,
but we deliberately avoided that complexity.

**What's next?**

We have a running cluster, but deploying updates currently means stopping each
node, unpacking the new tarball, and starting it again. For a trading bot, even
a few seconds of downtime means missed trades. In the next chapter, we'll use
OTP's hot code reloading to upgrade a running node without stopping it - push
new code, generate an appup describing how to transition each module's state,
and install it on the live system while trades continue to execute. The release
structure we built in this chapter is what makes that possible.

[Note] Please remember to run `mix format` to keep things nice and tidy.

The source code for this chapter can be found in the book's source code repository
(branch:
[chapter_24](https://github.com/Cinderella-Man/hands-on-elixir-and-otp-cryptocurrency-trading-bot-source-code/tree/chapter_24)).
